### 2018-02-01 stats/mcc_pilot.Rmd
Adding revised version of informativeness and anonymized stats

---
title: "MCC_pilot"
author: "Manuel Bohn"
date: "26 1 2018"
output: html_document
---

```{r setup}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)

## importing data
d <- bind_rows(read_csv(file="pref.data.csv"),
               read_csv(file="novel.data.csv")) %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"))

```

Sanity checks.
```{r sanity checks}

## sanity checks
d %>%
  group_by(trial_type, condition) %>%
  summarise(n = length(unique(id)), 
            correct = mean(correct)) %>%
  knitr::kable(digits = 2)

```
Plot condition and speaker change with 95%CI
```{r plot}

ms <- d %>%
  filter(trial_type == "test") %>%
  group_by(change, condition, id) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(ms, 
       aes(x = change, y = mean, fill = change, frame = condition)) +
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  facet_wrap(~ condition,
             labeller = as_labeller(c(`novelty`="Novelty", `preference`="Preference")))+ 
  theme_few(base_size = 12) + 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_fill_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

Conditions by trial.
```{r plot trials}
ms <- d %>%
  filter(trial_type == "test") %>%
  group_by(change, condition, trial, id) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(ms, 
       aes(x = trial, y = mean, col = change)) +
  geom_line(aes(group= change)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  facet_wrap(~ condition) + 
  theme_few() + 
  scale_colour_solarized()
```
Conditions by agents. Unequal distribution of agents across condition and change combinations leads to seemingly strange patterns for some agents. However, these strange results are always those with very few data points (e.g. Tiger novelty speaker change is only one data point). For those cells with more than five data points, the pattern is very similar to the overall pattern. 
```{r trials per condition and agent}
d %>%
  filter(trial_type == "test") %>%
  mutate(choiceAgent = ifelse( change == "false", agent, altAgent)) %>%
  group_by(choiceAgent,change, condition) %>%
  summarise(n = length(choiceAgent), 
            correct = mean(correct)) %>%
  knitr::kable(digits = 2)
```

```{r plot agents}
ms <- d %>%
  filter(trial_type == "test") %>%
  mutate(choiceAgent = ifelse( change == "false", agent, altAgent)) %>%
  group_by(change, condition, choiceAgent, id) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(ms, 
       aes(x = choiceAgent, y = mean, col = change)) +
  geom_line(aes(group= change)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  facet_wrap(~ condition) + 
  theme_few() + 
  scale_colour_solarized()
```

T-tests against chance for each condition*change combination
```{r}
## t tests against chance
d %>%
  filter(trial_type == "test") %>%
  group_by(change, condition, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(change,condition) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(condition,change,df,t_value,p_value) %>%
  knitr::kable(digits = 2)

```

Model testing for interaction between change and condition. No interaction between change and condition, and no effect of condition suggesting that speaker change manipualtion is similar in novelty and preference.
```{r}
## model
library(lme4)

data_cg <- d%>%
  filter(trial_type == "test")%>%
  mutate(choiceAgent = ifelse( change == "false", agent, altAgent)) %>%
  mutate(trial = scale(as.numeric(trial), center = TRUE, scale=FALSE))

lm <- glmer(correct ~ change * condition + trial +
              (change + trial|id) + (1|choiceAgent), 
            data = data_cg, family = binomial, control=glmerControl(optimizer="bobyqa",
                            optCtrl=list(maxfun=2e5)))

summary(lm)

```