### 2018-03-09 stats/simulations/mcc_kids_simulations.Rmd
kids data sim

---
title: "MCC Simulations"
author: "Manuel Bohn"
date: "10 2 2018"
output: html_document
---

```{r setup}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(ggpubr)
library(lme4)

```

```{r}
conditions <- read.csv("kids_condition_means.csv", sep=";")

generate_data <- function(condition_df, n_subs = 16, n_trials = 3) {
  tibble(count = condition_df$count,
         condition = condition_df$condition,
         age = condition_df$age,
         type = condition_df$type,
         correct = rbinom( n_subs * n_trials, 1,
                          condition_df$correct),
         subid = rep(1:n_subs, n_trials))
}

participant_data <- conditions %>%
  mutate(index = 1:n()) %>%
  split(.$index) %>%
  map_df(generate_data)%>%
  mutate(subid=paste(subid,age,condition,sep="_"))%>%
  arrange(condition, subid)

 participant_data %>%
  group_by(condition,age,type,subid) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct") %>%
  mutate(ciw = ci_upper-ci_lower) %>%
  group_by() %>%
  summarise(ciw = mean(ciw))%>%
  knitr::kable(digits = 3)

```

```{r}

ms <- participant_data %>%
  group_by(condition,age,type,subid) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(ms, 
       aes(x = type, y = mean, fill = type)) +
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  facet_grid(age~ condition)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few(base_size = 12) + 
  ylim(0,1)+
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())
```

```{r}

#
md <- participant_data %>%
  filter(condition == "novel")

lm <- glmer(correct ~ age * type + (1|subid), 
      data = md,
      family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(lm)
drop1(lm, test = "Chisq")

```