### 2018-03-10 stats/simulations/mcc_kids_simulations.Rmd
kids data sim

---
title: "MCC Simulations"
author: "Manuel Bohn"
date: "10 2 2018"
output: html_document
---

```{r setup}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(ggpubr)
library(lme4)

```

```{r}

### condition means based on adult data with slightly smaller effect for younger kids

conditions <- tibble (
  age = c(rep(c(2,2,4,4),3)),
  condition = c(rep(c("pref"),4),rep(c("novel"),4),rep(c("inf"),4)),
  type = c(rep(c("test","ctrl"),6)),
  correct = c(0.8666667,0.5416667,0.9666667,0.6416667,0.7333333,0.4916667,0.8333333,0.5916667,0.6716667,0.5,0.7416667,0.4583333)
)

generate_data <- function(condition_df, n_subs = 16, n_trials = 3) {
  tibble(condition = condition_df$condition,
         age = condition_df$age,
         type = condition_df$type,
         correct = rbinom( n_subs * n_trials, 1,
                          condition_df$correct),
         subid = rep(1:n_subs, n_trials))
}

### Average CI width across all condition, type and agegroup combinations.
### For adults (n = 40), this average CI width 0.172
### Average CI width as function of n kids: 12 = 0.261; 16 = 0.231; 24 = 0.192; 30 = 0.174; 40 = 0.150

for (i in 1:100) {
x[i] = conditions %>%
  mutate(index = 1:n()) %>%
  split(.$index) %>%
  map_df(generate_data)%>%
  mutate(subid=paste(subid,age,condition,sep="_"))%>%
  arrange(condition, subid) %>%
  group_by(condition,age,type,subid) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct") %>%
  mutate(ciw = ci_upper-ci_lower) %>%
  group_by() %>%
  summarise(ciw = mean(ciw))%>%
  pull()
}

mean(unlist(x))

```

```{r}

participant_data <- conditions %>%
  mutate(index = 1:n()) %>%
  split(.$index) %>%
  map_df(generate_data)%>%
  mutate(subid=paste(subid,age,condition,sep="_"))

ms <- participant_data %>%
  group_by(condition,age,type,subid) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(ms, 
       aes(x = type, y = mean, fill = type)) +
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  facet_grid(age~ condition)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few(base_size = 12) + 
  ylim(0,1)+
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())
```

```{r}

#
md <- participant_data %>%
  filter(condition == "novel")

lm <- glmer(correct ~ age * type + (1|subid), 
      data = md,
      family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(lm)

```