### 2018-04-12 models/predictions.Rmd
updating Ex3 and model and predictions

---
title: "MCC Predictions"
author: "Manuel Bohn"
date: "9 4 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(ggthemes)
```

# Modeling Informativeness

Speaker optimality parameter of 2.25 produces model predictions that match out empirical results. The following predictions therefore use this parameter.

Prior = measured proportion this type of object is being chosen.

## "Dampening"" model

### Preference inconcgruent
ObjectPrior = Categorical({vs: all_objects, ps: [0.9666667, 0.9666667, 0.0333333] })

### Preference congruent
ObjectPrior = Categorical({vs: all_objects, ps: [0.0333333, 0.0333333, 0.9666667] })

### Novelty inconcgruent
var ObjectPrior = Categorical({vs: all_objects, ps: [0.8333333, 0.8333333, 0.1666667] })

### Novelty concgruent
var ObjectPrior = Categorical({vs: all_objects, ps: [0.1666667, 0.1666667, 0.8333333] })

```{r , echo=FALSE}

pred <- data.frame(
  Experiment = c("Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty","Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty","Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty"),
  Change = c("Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker"),
  Alignment = c("Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent"),
  Model = c("Full model","Full model","Full model","Full model","Full model","Full model","Full model","Full model","No prior","No prior","No prior","No prior","No prior","No prior","No prior","No prior", "Prior only","Prior only","Prior only","Prior only","Prior only","Prior only","Prior only","Prior only"),
  corr = c(0.9999672744165804,0.03460862921907267,0.9083230583454356,0.5090645649737232,0.9930727540990627,0.20045767964927574,0.8603373265403738,0.594975610001643,0.742,0.742,0.742,0.742,0.742,0.742,0.742,0.742,0.967,0.033,0.642,0.358,0.833,0.167,0.592,0.408))

pred %>%
  knitr::kable(digits = 3)

ggplot(pred, 
       aes(x = Model, y = corr, fill = Model)) +
  geom_bar(stat="identity") + 
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x=" ",y="Proportion Choosing More Informative")+
  facet_grid(Experiment ~ Change*Alignment)+
  theme_few(base_size = 12) + 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.x=element_blank())+
  ylim(0,1)

```

At what ratio does it flip back to more informative?

```{r , echo=FALSE}

pred <- data.frame(
  Prior = c("90/10","80/20","70/30","60/40","50/50"),
  corr = c(0.1118794400899893,0.24897738407515466,0.40911964439987003,0.5807810413750243,0.74)
)

ggplot(pred, 
       aes(x = Prior, y = corr, fill = Prior)) +
  geom_bar(stat="identity") + 
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Choosing More Informative")+
  theme_few(base_size = 12) +
  scale_fill_discrete(name = "Prior: less inf / more inf")+
  ylim(0,1)+
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())

```