### 2018-04-27 stats/mcc_pref.weak.Rmd
novelty weak 2 second exp

---
title: "MCC_Preference_Weak"
author: "Manuel Bohn"
date: "26 4 2018"
output: html_document
---

```{r setup}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)

## importing data
d <- read_csv(file="pref.weak.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"))

```

Sanity checks.
```{r sanity checks}

## sanity checks
d %>%
  group_by(trial_type) %>%
  summarise(n = length(unique(id)), 
            correct = mean(correct)) %>%
  knitr::kable(digits = 2)

# check if someone needs to be excluded because wrong in training
# no one
d %>%
  filter(trial_type == "train") %>%
  group_by(id)%>%
  summarise(correct = mean(correct)) %>%
  filter(correct == 0)

```

Plot for speaker change with 95%CI
```{r plot}

p1 <- d %>%
  filter(trial_type == "test") %>%
  group_by(change, id) %>%
  summarise(correct = mean(correct))

p2 <- p1 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p1, aes(x = change, y = correct, col = change, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = change, y = mean, col = change, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

T-tests against chance for each speaker change conditions
```{r}
## t tests against chance
d %>%
  filter(trial_type == "test") %>%
  group_by(change, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(change) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(change,df,t_value,p_value) %>%
  knitr::kable(digits = 2)

```

ChoiceAgent can be seen as an approximation for "item".

```{r}
## model
library(lme4)

data_cg <- d%>%
  filter(trial_type == "test")%>%
  mutate(choiceAgent = ifelse( change == "false", agent, altAgent)) %>%
  mutate(trial = scale(as.numeric(trial), center = TRUE, scale=FALSE))

# maximal converging model
lm <- glmer(correct ~ change + (change |id) + (change |choiceAgent), 
              data = data_cg, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(lm)
```

Conditions by trial.
```{r plot trials}
ms <- d %>%
  filter(trial_type == "test") %>%
  group_by(change, trial, id) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(ms, 
       aes(x = trial, y = mean, col = change)) +
  geom_line(aes(group= change)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  theme_few() + 
  scale_colour_solarized()
```

Conditions by agents. Unequal distribution of agents across change leads to seemingly strange patterns for some agents. However, these strange results are always those with very few data points (e.g. Tiger novelty speaker change is only one data point). For those cells with more than five data points, the pattern is very similar to the overall pattern. Agents are ranodmly selected and assigned to speaker change in the beginning of the experiment. This causes the unequal distribution.
```{r trials per condition and agent}
d %>%
  filter(trial_type == "test") %>%
  mutate(choiceAgent = ifelse( change == "false", agent, altAgent)) %>%
  group_by(choiceAgent,change) %>%
  summarise(n = length(choiceAgent), 
            correct = mean(correct)) %>%
  knitr::kable(digits = 2)
```

```{r plot agents}
ms <- d %>%
  filter(trial_type == "test") %>%
  mutate(choiceAgent = ifelse( change == "false", agent, altAgent)) %>%
  group_by(change, choiceAgent, id) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(ms, 
       aes(x = choiceAgent, y = mean, col = change)) +
  geom_line(aes(group= change)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  theme_few() + 
  scale_colour_solarized()
```