### 2018-05-10 stats/mcc_priors_overview.Rmd
add weaker prior dis-preference

---
title: "MCC Prior Overview"
author: "Manuel Bohn"
date: "9 4 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
```

```{r data}
d3 <- read_csv(file="novel.weak.data2.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "2_medium")

d4 <- read_csv(file="novel.weak.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "1_weak")

d1 <- read_csv(file="novel.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "3_strong")

d2 <-read_csv(file="pref.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "3_strong")

d5 <-read_csv(file="pref.weak.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "2_medium")

d6 <-read_csv(file="pref.weak.data3.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "1_weak")

d <- bind_rows(d1,d2,d3,d4,d5,d6)

# check if someone got both training trials wrong
d %>%
  filter(trial_type == "train") %>%
  group_by(id)%>%
  summarise(sum = sum(correct)) %>%
  filter(sum == 0)

# check if someone did multiple experiments
# yes, six people
d %>%
  filter(trial_type == "train") %>%
  group_by(id)%>%
  summarise(n = length(correct)) %>%
  filter(n > 2)

# exclude those who did both experiments
# keep first time they participated

d <- d %>% 
  distinct(id, alltrial, .keep_all = TRUE)
```

```{r plot data}

p1 <- d %>%
  filter(trial_type == "test") %>%
  group_by(strength,condition,change, id) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(p1, 
       aes(x = change, y = mean, fill = change)) +
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few(base_size = 12) + 
  facet_grid(condition ~ strength , scales = "free_x", space = "free_x" ) +
  ylim(0,1)+
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_fill_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))

```