### 2018-10-29 stats/mcc_summary_inf_pilot_kids.Rmd

---
title: "MCC Informativeness - summary pilots kids"
author: "Manuel Bohn"
date: "7 2 2018"
output: html_document
---

```{r setup, include = FALSE}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(jsonlite)

## round 1 adult design
d_round_1 <- read_csv(file="kids_inf.data.csv")%>%
  filter(trial != "train", control == "false")%>%
  mutate(setup = "table",
         move = "no_move",
         turn = "no_turn",
         wording = "there_is_a_dax",
         filler = "no_filler",
         choice = "object",
         subage = age,
         experiment = "1_round1",
         subage = as.character(subage))%>%
  select(subid,subage,setup,control,trial,move,turn,wording,filler,choice,experiment,agent,correct)

## introducing filler trials
files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/filler")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/filler/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$data$data)
  raw_data <- bind_rows(raw_data, id)
}

d_filler <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "train3",
         trial != "train4")%>%
  mutate(setup = "table",
         move = "no_move",
         turn = "no_turn",
         wording = "there_is_a_dax",
         filler = "filler",
         choice = "object",
         experiment = "2_filler",
         subage = as.character(subage))%>%
  select(subid,subage,setup,control,trial,move,turn,wording,filler,choice,experiment,agent,correct)

## moving agent
files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/moving_agent")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/moving_agent/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$data$data)
  raw_data <- bind_rows(raw_data, id)
}

d_move_agent <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "train3",
         trial != "train4")%>%
  mutate(setup = "table",
         move = "move",
         turn = "no_turn",
         wording = "there_is_a_dax",
         filler = "no_filler",
         choice = "object",
         experiment = "3_move_agent",
         subage = as.character(subage))%>%
  select(subid,subage,setup,control,trial,move,turn,wording,filler,choice,experiment,agent,correct)%>%
  mutate(subage = ifelse(subage == "3 ", "3", subage))

## moving agent, turning before starting to move
files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/moving_turning")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/moving_turning/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$data$data)
  raw_data <- bind_rows(raw_data, id)
}

d_move_turn <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "train3",
         trial != "train4")%>%
  mutate(setup = "table",
         move = "move",
         turn = "turn",
         wording = "there_is_a_dax",
         filler = "no_filler",
         choice = "object",
         experiment = "4_move_turn",
         subage = as.character(subage))%>%
  select(subid,subage,setup,control,trial,move,turn,wording,filler,choice,experiment,agent,correct)

## introduce new wording, spoken by experimenter
files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/new_wording")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/new_wording/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$data$data)
  raw_data <- bind_rows(raw_data, id)
}

d_new_wording <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "filler1",
         trial != "filler2")%>%
  mutate(setup = "table",
         move = "move",
         turn = "turn",
         wording = "Here_is_a_table_with_a_dax",
         filler = "filler",
         choice = "object",
         experiment = "5_new_wording",
         subage = as.character(subage))%>%
  select(subid,subage,setup,control,trial,move,turn,wording,filler,choice,experiment,agent,correct)%>%
  filter(subid != "Test")

## introducing robots, agent is looking, experimenter reads text
files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/robots_look_1")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/robots_look_1/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$data$data)%>%
    mutate(subid = jd$data$subid,
           subage = jd$data$subage)
  raw_data <- bind_rows(raw_data, id)
}

d_robot_look <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "filler1",
         trial != "filler2")%>%
  mutate(setup = "robot",
         move = "no_move",
         turn = "no_turn",
         wording = "Here_is_a_table_with_a_dax",
         filler = "filler",
         choice = "object",
         experiment = "6_robot_look",
         subage = as.character(subage))%>%
  select(subid,subage,setup,control,trial,move,turn,wording,filler,choice,experiment,agent,correct)%>%
  filter(subid != "Test")

## robots, agent is pointing
files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/robots_point_1")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/robots_point_1/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$data$data)
  raw_data <- bind_rows(raw_data, id)
}

d_robot_point <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "filler1",
         trial != "filler2")%>%
  mutate(setup = "robot",
         move = "no_move",
         turn = "no_turn",
         wording = "Here_is_a_table_with_a_dax",
         filler = "filler",
         choice = "object",
         experiment = "7_robot_point",
         subage = as.character(subage))%>%
  select(subid,subage,setup,control,trial,move,turn,wording,filler,choice,experiment,agent,correct)

## robots talk themselves
files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/robots_point_3_sound")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/robots_point_3_sound/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$data$data)
  raw_data <- bind_rows(raw_data, id)
}

d_robot_talk <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "filler1",
         trial != "filler2")%>%
  mutate(setup = "robot",
         move = "no_move",
         turn = "no_turn",
         wording = "Here_is_a_table_with_a_dax",
         filler = "filler",
         choice = "object",
         experiment = "8_robot_point_talk",
         subage = as.character(subage))%>%
  select(subid,subage,setup,control,trial,move,turn,wording,filler,choice,experiment,agent,correct)

## robots map onto container
files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/robots_point_4_map")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/robots_point_4_map/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$data$data)
  raw_data <- bind_rows(raw_data, id)
}

d_robot_map_container <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "filler1",
         trial != "filler2")%>%
  mutate(setup = "robot",
         move = "no_move",
         turn = "no_turn",
         wording = "Here_is_a_table_with_a_dax",
         filler = "filler",
         choice = "container",
         experiment = "9_robot_map_container",
         subage = as.character(subage))%>%
  select(subid,subage,setup,control,trial,move,turn,wording,filler,choice,experiment,agent,correct)

pilot_data <- bind_rows(
  d_round_1,
  d_filler,
  d_move_agent,
  d_move_turn,
  d_new_wording,
  d_robot_look,
  d_robot_point,
  d_robot_talk,
  d_robot_map_container
)

```

# Descriptives

```{r}

```

# Overview by experiment

## Experiment descriptions

* 1_round1: Table setup (like adults) including control trials (4 objects)
* 2_filler: Table setup without control trials but with filler trials
* 3_move_agent: Table setup with agent moving in front of table before asking
* 4_move_turn: Table setup with agent turning to the other table before moving to the target table and requesting
* 5_new_wording: Table setup with experimenter reading out prompts. Wording changed to "Here is a table with a dax".
* 6_robot_look: Robot setup with experimenter reading prompts. Agent looks to robot when making request. Wording (from here onwards): "This is a robot with a dax".
* 7_robot_point: Robot setup, experimenter reading prompts. Agent points twice while labeling and before requesting.
* 8_robot_point_talk: Robot setup, recorded request. Pointing twice to target robot before requesting.
* 9_robot_map_container: Robot setup, recorded request. Pointing twice during labeling. Children pick container instead of object directly.

```{r plot, echo = FALSE, warning = FALSE, fig.width=12}

p <- pilot_data %>%
  group_by(experiment,subage, subid) %>%
  summarise(correct = mean(correct))

p1 <- p %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p, aes(x = 1, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p1, aes(x = 1, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_grid(subage~experiment)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())
```

```{r plot, echo = FALSE, warning = FALSE}

p <- pilot_data %>%
  filter(subage != "2")%>%
  group_by(experiment, subid) %>%
  summarise(correct = mean(correct))

p1 <- p %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p, aes(x = 1, y = correct, col = experiment, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p1, aes(x = 1, y = mean, col = experiment, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_grid(.~experiment)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())
```

# Model across experiments
```{r model, cache = TRUE, warning = FALSE}
library(lme4)

lm <- glmer(correct ~ subage * setup + subage * move + subage * turn + subage * wording + subage * filler + subage * choice + (subage  | agent) + (subage | experiment) + (1 | subid), 
              data = pilot_data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(lm)

```

# Indivudal variations

## Setup

```{r, echo=FALSE}
p <- pilot_data %>%
  group_by(setup,subage, subid) %>%
  summarise(correct = mean(correct))

p1 <- p %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p, aes(x = 1, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p1, aes(x = 1, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_grid(setup~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())
```

## Wording

```{r, echo=FALSE}
p <- pilot_data %>%
  group_by(wording,subage, subid) %>%
  summarise(correct = mean(correct))

p1 <- p %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p, aes(x = 1, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p1, aes(x = 1, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_grid(wording~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())
```

## Move (only table setup)

```{r, echo=FALSE}
p <- pilot_data %>%
  filter(setup == "table")%>%
  group_by(move,subage, subid) %>%
  summarise(correct = mean(correct))

p1 <- p %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p, aes(x = 1, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p1, aes(x = 1, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_grid(move~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())
```

## Filler

```{r, echo=FALSE}
p <- pilot_data %>%
  group_by(filler,subage, subid) %>%
  summarise(correct = mean(correct))

p1 <- p %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p, aes(x = 1, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p1, aes(x = 1, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_grid(filler~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())
```