### 2018-11-23 stats/kids_model_predictions.Rmd

---
title: "MCC Kids Model Predictions"
author: "Manuel Bohn"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
library(coda)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}

```
# Pretest

Loading data

```{r pretest data}

pref_data <- read_csv(file="data/kids_pref_data.csv")%>%
  filter(age != "2",
         trial_type != "train")%>%
  mutate(age_bin = factor(age),
         experiment = "preference_pretest",
         condition = ifelse(change == "false", "same_speaker", "different_speaker"),
         agent = ifelse( change == "false", agent, altAgent))%>%
  select(experiment,subid,age_bin,age_num,trial,condition,agent,correct)

inf_data <- read_csv(file="data/kids_inf_data.csv")%>%
  filter(subage != "2",
         trial != "filler1",
         trial != "filler2")%>%
  mutate(age_bin = factor(subage),
         experiment = "informativeness_pretest",
         condition = ifelse(control == "no", "test", "control"))%>%
  select(experiment,subid,age_bin,age_num,trial,condition,agent,correct)

pre_data <- bind_rows(pref_data,
                      inf_data)

```

# Pretest results

```{r pretest plot}

pre_p1 <- pre_data %>%
  group_by(experiment,condition,age_num,subid)%>%
  summarise(correct = mean(correct))

ggplot(data = pre_p1, aes(x = age_num, y = correct, col = condition)) +
  geom_smooth(method = "lm", se = F)+
  geom_jitter(alpha = .5, height = 0, width = 0)+
  geom_hline(yintercept = 0.5, lty=2)+
  facet_wrap(~experiment)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized()

```

# Model predictions

## Pretest slopes
```{r rsaUtils, include = FALSE}
rsaUtils <- '
var all_objects = [
{ shape: "triangle", id:1, location: 1},  
{ shape: "triangle", id:2, location: 2},
{ shape: "circle", id:1, location: 2}
]

var labels = ["dax","wug"]

var lexicon1 = function(utterance, obj){
utterance.label == "dax" ? obj.shape == "triangle" :
utterance.label == "wug" ? obj.shape == "circle" : 
true
}

var lexicon2 = function(utterance, obj){
utterance.label == "dax" ? obj.shape == "circle" :
utterance.label == "wug" ? obj.shape == "triangle" : 
true
}

var lexiconObjects = {
"dax = triangle": {
triangle: "dax", circle: "wug"
},
"dax = circle": {
triangle: "wug", circle: "dax"
},
}

var lexiconObject = {
"dax = triangle": lexicon1,
"dax = circle" : lexicon2
}

var point = function(utterance, obj){
return obj.location == utterance.point
}

var utterancePrior = function(obj, lexiconName){
var locationsWithShape = _.map(_.filter(all_objects, {shape: obj.shape}), "location")
var point = uniformDraw(locationsWithShape)
var label = lexiconObjects[lexiconName][obj.shape]
return {label: label, point: point}
}

var LexiconPrior = Categorical({vs: ["dax = triangle","dax = circle" ], ps: [1, 1]})
'
```

```{r rsa model, include = FALSE}

rsaModel <- '

var pragmaticListener = function(utterance, priorProbs){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior);
var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
var S1 = speaker(obj, lexiconName, priorProbs);
observe(S1, utterance)
return obj.shape == "circle" ? 1 : 0
}})
}

var speaker = function(obj, lexiconName, priorProbs){
Infer({method: "enumerate", model: function(){
var utterance = utterancePrior(obj, lexiconName);
var L0 = literalListener(utterance, priorProbs);
 factor(speakerOptimality * L0.score(obj.shape))
return utterance
}})
}

var literalListener = function(utterance, priorProbs){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior); 
var lexicon = lexiconObject[lexiconName];
var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
if ("label" in utterance) {
 var truthValue = lexicon(utterance, obj);
 condition(truthValue)
}
if (utterance.point) {
 var truthValuePoint = point(utterance, obj);
 condition(truthValuePoint)
}
return obj.shape 
}})
}

var addNoise = function(dist, noiseParam){
   Infer({model: function(){ 
      return flip(noiseParam) ? uniformDraw([0, 1]) : sample(dist)
    }
   })
}
'
```

```{r speaker optimality slope}
soSlopeWebppl <- '

var infData = dataFromR.infData
var priorProbs = [.5, .5, .5]

var model  = function(){

  var so_slope = uniform(-5, 5)
  var so_int = uniform(0, 20)

    map(function(age){

      var speakerOptimality = so_int  + so_slope * (age - infData.minage)
      var rsaPredictions = pragmaticListener({label: "dax", point: 2 }, priorProbs)

      observe(rsaPredictions, infData.correct)

      return  [infData.experiment + "_" + infData.age_num, Math.exp(model.score(1))]

    }, infData.age_num)

}
'
```

```{r}
soSlope <- webppl(
  program_code = paste(rsaUtils, rsaModel, soSlopeWebppl , sep='\n'),
  data = list(infData = inf_data%>%mutate(minage = min(age_num))), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 2000, burn = 500, verbose = T)
)
```

```{r}
save.image(file="kids_ex3_predictions.RData")
```