### 2019-01-09 stats/kids/kids_Ex3_pref.Rmd

---
title: "MCC Ex3 Preference Kids"
author: "Manuel Bohn"
output: html_document
---

```{r import data, include = F}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(jsonlite)
library(readxl)

## importing data
files <- dir("~/Work/MCC/git-mcc/kids_EX3/pref/")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_EX3/pref/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,42,str_length(jf)-11)
  id <- as_data_frame(jd$data$data) %>% mutate(test_date = date)
  raw_data <- bind_rows(raw_data, id)
}

raw_data <- raw_data %>%
  mutate(subid = ifelse(subid == "181128_3_ex3p....","181128_3_ex3p" , subid),
         subid = ifelse(subid == "181128_1_ex3p.","181128_1_ex3p" , subid),
         subid = ifelse(subid == " 181218_10_ex3p","181218_10_ex3p" , subid))

raw_data%>%
  #group_by(paste(subid,test_date))%>%
  group_by(subid)%>%
  summarise(n= n())

unique(raw_data$subid)

log <- read_excel("../../../MCC-subject_log.xlsx", 1)%>%
  filter(Condition == "ex3p")%>%
  select(subid,experimenter,keep_drop,sex,dob)

#unique(log$subid)
sum(log$keep_drop == "drop")

d <- left_join(raw_data,log, by = "subid")%>%
  filter(keep_drop == "keep")%>%
  mutate(dob = as.numeric(dob),
         dob = as.Date(dob,origin = "1899-12-30"),
         age_num = lubridate::time_length(difftime(test_date,dob), "years"),
         check_age = ifelse(substr(age_num,1,1) == subage, T, F))

unique(d$check_age)

```

# Sample size per age bin

```{r kids per age group,echo = F }

d %>%
  group_by(subage,subid)%>%
  summarise(n_trials = n())%>%
  mutate(full_data = ifelse(n_trials == 10,"yes","no"))%>%
  group_by(subage)%>%
  summarise(n = length(unique(subid)),
            n_full_data = sum(full_data == "yes"))%>%
  knitr::kable(digits = 2)

```

# Datapoints per cell

```{r kids per age group,echo = F }

d %>%
  filter(trial != "filler1",trial != "filler2")%>%
  group_by(subage,speaker,alignment)%>%
  summarise(datapoints = n())%>%
  knitr::kable(digits = 2)

```

# Plot filler by age

```{r plot by age, echo= FALSE}
pf1 <- d %>%
      filter(trial == "filler1" | trial == "filler2")%>%
  group_by(subage, subid) %>%
  summarise(correct = mean(correct_inf))

pf2 <- pf1 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = pf1, aes(x = subage, y = correct, col = subage, alpha = .2),width = .1,height = .025)+
  geom_pointrange(data = pf2, aes(x = subage, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())
```

# Plot by age group

```{r plot by age, echo= FALSE}

p1 <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by(speaker, alignment ,subage, subid) %>%
  summarise(correct = mean(correct_inf))

p2 <- p1 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p1, aes(x = alignment, y = correct, col = alignment, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = alignment, y = mean, col = alignment, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_grid(speaker~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

# Plot by age continous

```{r plot by age, echo= FALSE}

p3 <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by(speaker, alignment ,age_num, subid) %>%
  summarise(correct = mean(correct_inf))

ggplot(data = p3, aes(x = age_num, y = correct_inf)) +
  geom_jitter(width = .025,height = .025, alpha = .2)+
  geom_smooth(method = "glm", method.args = list(family = "binomial"),  se = T,  alpha = .5, size = .4)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="Age",y="Proportion Expected Choice")+
  theme_few() +
  facet_grid(speaker~alignment)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  coord_fixed(ratio = 2)

```

```{r model}

library(lme4)

dm <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  mutate(age_num = scale(age_num, center = T, scale = T))

mex3 <- glmer(correct_inf ~ age_num*speaker*alignment 
      + (speaker*alignment | subid) 
      + (age_num*speaker*alignment | agent), 
      family = "binomial",
      data = dm,
      control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(mex3)
```

```{r chance_comp}
mod <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by(subage,speaker,alignment) %>%
  do(models = glmer(correct_inf ~ 1+ (1|subid),data=., family=binomial, control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))))

library(broom)

tidy(mod, models) %>%
  filter(term =="(Intercept)")
```