### 2019-02-08 stats/kids/kids_model_predictions.Rmd
update kids model preictions

---
title: "MCC Kids Model Predictions"
author: "Manuel Bohn"
output: 
  html_document:
      code_folding: hide
---

```{r, include = F}
knitr::opts_chunk$set(echo = TRUE, warning = F)
```

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(ggthemes)
library(ggpubr)
library(langcog)
library(rwebppl)
library(coda)
library(reshape2)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}

```

# Pretest

```{r pretest data, include=FALSE}

pref_data <- read_csv(file="../data/kids_pref_data.csv")%>%
  filter(age != "2",
         trial_type != "train")%>%
  mutate(age_bin = factor(age),
         experiment = "preference_pretest",
         condition = ifelse(change == F | change == "false", "same_speaker", "different_speaker"),
         agent = ifelse( change == F | change == "false", agent, altAgent),
         minage = min(age_num))%>%
  select(experiment,subid,age_bin,age_num,trial,condition,agent,correct,minage)

inf_data <- read_csv(file="../data/kids_inf_data.csv")%>%
  filter(subage != "2",
         trial != "filler1",
         trial != "filler2")%>%
  mutate(age_bin = factor(subage),
         experiment = "informativeness_pretest",
         condition = ifelse(control == "no", "test", "control"),
        minage = min(age_num))%>%
  select(experiment,subid,age_bin,age_num,trial,condition,agent,correct,minage)

pre_data <- bind_rows(pref_data,
                      inf_data)

ex3_data <- read_csv(file="../data/kids_ex3_data.csv")%>%
  filter(trial != "filler1",trial != "filler2")

```

## Pretest results

This is the data we use to estimate the speaker optimality nad the age specific prior.

```{r pretest plot so, echo = F}

pre_so <- pre_data %>%
  filter(experiment == "informativeness_pretest")%>%
  group_by(age_num,subid)%>%
  summarise(correct = mean(correct))

plot_inf <- ggplot(data = pre_so, aes(x = age_num, y = correct)) +
  geom_jitter(alpha = 1, height = 0.1, width = 0.1, col = "grey")+
  geom_smooth(method = "lm", se = T, col = "black",fill = "darkgreen", alpha = .4, size = .4)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="age",y="Proportion Expected Choice")+
  theme_few() +
  ggtitle("Informativeness data")+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  coord_fixed(ratio = 2)
```

```{r pretest plot prior, echo = F}

pre_prior <- pre_data %>%
  filter(experiment == "preference_pretest")%>%
  group_by(condition,age_num,subid)%>%
  summarise(correct = mean(correct))

plot_pref <- ggplot(data = pre_prior, aes(x = age_num, y = correct, col = condition)) +
  geom_jitter(alpha = 0.5, height = 0.1, width = 0.1)+
  geom_smooth(aes(fill = condition),method = "lm", se = T, col = "black",  alpha = .5, size = .4)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ggtitle("Preference data")+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  coord_fixed(ratio = 2)+
  scale_color_solarized()+
  scale_fill_solarized()

```

```{r, echo = F, fig.width= 8}
ggarrange(plot_inf, plot_pref, widths = c(1.29,2))
```

## RSA model

### Model utensils.

There are two types of objects, one used in the 3-object world of the RSA model (`all_objects`) and one used in the 2-object world of the prior (`prior_objects`).

```{r rsaUtils}
rsaUtils <- '
var all_objects = [
{ shape: "triangle", id:1, location: 1},  
{ shape: "triangle", id:2, location: 2},
{ shape: "circle", id:1, location: 2}
]

var prior_objects = [
{ shape: "triangle", id:1, location: 1},  
{ shape: "circle", id:1, location: 2}
]

var labels = ["dax","wug"]

var lexicon1 = function(utterance, obj){
utterance.label == "dax" ? obj.shape == "triangle" :
utterance.label == "wug" ? obj.shape == "circle" : 
true
}

var lexicon2 = function(utterance, obj){
utterance.label == "dax" ? obj.shape == "circle" :
utterance.label == "wug" ? obj.shape == "triangle" : 
true
}

var lexiconObjects = {
"dax = triangle": {
triangle: "dax", circle: "wug"
},
"dax = circle": {
triangle: "wug", circle: "dax"
},
}

var lexiconObject = {
"dax = triangle": lexicon1,
"dax = circle" : lexicon2
}

var point = function(utterance, obj){
return obj.location == utterance.point
}

var utterancePrior = function(obj, lexiconName){
var locationsWithShape = _.map(_.filter(all_objects, {shape: obj.shape}), "location")
var point = uniformDraw(locationsWithShape)
var label = lexiconObjects[lexiconName][obj.shape]
return {label: label, point: point}
}

var LexiconPrior = Categorical({vs: ["dax = triangle","dax = circle" ], ps: [1, 1]})
'
```

### RSA model code

```{r rsa model}

rsaModel <- '
var literalListener = cache(function(utterance, priorProbs){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior); 
var lexicon = lexiconObject[lexiconName];
var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
if ("label" in utterance) {
 var truthValue = lexicon(utterance, obj);
 condition(truthValue)
}
if (utterance.point) {
 var truthValuePoint = point(utterance, obj);
 condition(truthValuePoint)
}
return obj.shape 
}})
})

var speaker = cache(function(obj, lexiconName, priorProbs, speakerOptimality){
Infer({method: "enumerate", model: function(){
var utterance = utterancePrior(obj, lexiconName);
var L0 = literalListener(utterance, priorProbs);
 factor(speakerOptimality * L0.score(obj.shape))
return utterance
}})
})

var pragmaticListener = cache(function(utterance, priorProbs, speakerOptimality){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior);
var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
var S1 = speaker(obj, lexiconName, priorProbs, speakerOptimality);
observe(S1, utterance)
return obj.shape == "circle" ? 1 : 0
}})
})

var addNoise = function(dist, noiseParam){
   Infer({model: function(){ 
      return flip(noiseParam) ? uniformDraw([0, 1]) : sample(dist)
    }
   })
}
'
```

### Priors for intercept and slope for speaker optimality and preference

Sanity check for the priors on slope and intercept.

#### Speaker optimality

```{r prior speaker optimality slope}
priorSoWebppl <- '
var infData = dataFromR
var priorProbs = [.5, .5, .5] 
var speakerOptimality = []
var model  = function(){
  var so_slope = uniformDrift({a: -2, b: 2, width: 0.4})
  var so_int = uniformDrift({a: -2, b: 2, width: 0.4})
    map(function(row){
      var age = row.age_num
      var speakerOptimality = so_int  + so_slope * (age - row.minage)
      var rsaPredictions = pragmaticListener({label: "dax", point: 2 }, priorProbs, speakerOptimality)

      //display(JSON.stringify(row.minage))
    }, infData)
  return extend({so_int: so_int, so_slope: so_slope})
}
'
```

```{r speaker optimality prior pred}
so_prior <- readRDS("../saves/so_prior.rds")
# so_prior<- webppl(
#   program_code = paste(rsaUtils, rsaModel, priorSoWebppl , sep='\n'),
#   data = inf_data%>%mutate(minage = min(age_num)), 
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 1,
#   inference_opts = list(method = "forward", samples = 1000, verbose = T)
# )
#saveRDS(so_prior, "../saves/so_prior.rds")

priorPred <- so_prior%>%
  spread(Parameter, value)%>%
  mutate(chain = factor(Chain))

so_prior <- so_prior%>%
  mutate(chain = factor(Chain))

mean_so_pre <-priorPred%>%
  summarise(int = mean(so_int),
            slope = mean(so_slope))

so_prior_plot <- ggplot(data = priorPred) +
  geom_abline(aes(intercept = so_int, slope = so_slope), col = "grey", alpha = 0.3)+
  geom_abline(data = mean_so_pre, aes(intercept = int, slope = slope), size = 0.4)+
  scale_x_continuous(limits=c(0, 2), name="age", breaks = c(0:2), labels = c(3:5)) +
  scale_y_continuous(limits=c(-4, 4), name="Speaker optimality")+
  ggtitle("Model speaker optimality prior")+
  coord_fixed(ratio=1/4)+
  theme_few()
```

#### Preference

```{r prior preference congruent}
priorPreWebppl <- '
var prefData = dataFromR
var priorProbs = [] 
var logistic = function(x) {1 / (1 + Math.exp(-x))}
var model  = function(){
  var pref_slope = uniformDrift({a: -2, b: 2, width: 0.4})
  var pref_int = uniformDrift({a: -2, b: 2, width: 0.4})
    map(function(row){
      var age = row.age_num
      var priorReg = logistic(pref_int + pref_slope * (age - row.minage))
      var priorProbs= [1-priorReg, priorReg]
      var priorPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: prior_objects, ps: priorProbs}));
      return obj.shape == "circle" ? 1 : 0
      }})

      // observe(priorPredictions, row.correct)
     // display(JSON.stringify(priorReg))
    }, prefData)
  return extend({pref_int: pref_int, pref_slope: pref_slope})
}
'
```

```{r}
pref_prior_con <- readRDS("../saves/pref_prior_con.rds")

# pref_prior_con<- webppl(
#   program_code = paste(rsaUtils, rsaModel, priorPrefConWebppl , sep='\n'),
#   data = pref_data%>%mutate(minage = min(age_num)), 
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 1,
#   inference_opts = list(method = "forward", samples = 1000, verbose = T)
# )
 # saveRDS(pref_prior_con, "../saves/pref_prior_con.rds")

prefPriorPredCon <- pref_prior_con%>%
  spread(Parameter, value)

pref_prior_con <- pref_prior_con%>%
  mutate(chain = factor(Chain),
         Alignment = "congruent")

x_plot <- seq(0, 2, by = 0.1)
model_predictions <- sapply(1:length(prefPriorPredCon$pref_int), function(idx) {
            plogis(prefPriorPredCon$pref_int[idx] + prefPriorPredCon$pref_slope[idx] * x_plot)
        })
colnames(model_predictions) <- 1:length(prefPriorPredCon$pref_int)
plot_pref_prior_con <- as.data.frame(cbind(x_plot, model_predictions))
plot_pref_prior_con <- melt(plot_pref_prior_con, id.vars = "x_plot", variable.name = "iteration", 
        value.name = "y_plot")

plot_prior_con_mean <- plot_pref_prior_con%>%
  group_by(x_plot)%>%
  summarise(y_plot = mean(y_plot))

pref_prior_plot <- ggplot(plot_pref_prior_con) + 
  geom_line(aes(x_plot, y_plot, group = iteration), col = "grey",alpha = 0.2) +
  #geom_ribbon(data = plot_prior_mean, aes(x =x_plot, ymin = ci_lower, ymax = ci_upper), fill = "red", alpha = 0.5) +
  geom_line(data = plot_prior_con_mean, aes(x_plot, y_plot), col = "black", size = 0.4) +
  xlab("age") + ylab("Proportion Expected Choice") +
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0:2), labels = c(3:5)) +
  scale_y_continuous(limits = c(0, 1), name="P in favor of preferred object")+
  ggtitle("Model preference prior")+
  coord_fixed(ratio=2)+
  theme_few()
```

```{r, echo = F, fig.width= 8}
ggarrange(so_prior_plot, pref_prior_plot)
```

There is no information in the priors. All structure in the models further down comes from the data

# Model Predictions

## Model parameters by age bin

```{r}
paramModelWebppl <- '

var levels = function(df, label){
  return _.uniq(_.map(df, label));
}
var prefData = _.filter(dataFromR, {experiment: "preference_pretest"})
var infData = _.filter(dataFromR, {experiment: "informativeness_pretest"})
// make finer grained as needed
var binned_age_range = _.range(0, 2, 0.05)
var all_conditions = levels(prefData, "condition")
var foreach = function(fn, lst) {
    var foreach_ = function(i) {
        if (i < lst.length) {
            fn(lst[i]);
            foreach_(i + 1);
        }
    };
    foreach_(0);
};
var logistic = function(x) {1 / (1 + Math.exp(-x))}
var model  = function(){
  var so_slope = uniformDrift({a: -2, b: 2, width: 0.4})
  var so_int = uniformDrift({a: -2, b: 2, width: 0.4})
  foreach(function(row){
    var age = row.age_num
    var speakerOptimality = so_int  + so_slope * (age - infData[0].minage)

    var inf_priorProbs = [.5, .5, .5] 
    var rsaPredictions = pragmaticListener({label: "dax", point: 2 },
    inf_priorProbs, speakerOptimality)

    observe(rsaPredictions, row.correct)
  }, infData)
  var pref_params = map(function(cndtn){
    var conditionData = _.filter(prefData, {condition: cndtn})

    var pref_slope = uniformDrift({a: -2, b: 2, width: 0.4})
    var pref_int = uniformDrift({a: -2, b: 2, width: 0.4})

    foreach(function(row){
      var age = row.age_num
      var priorReg = logistic(pref_int + pref_slope * (age - row.minage))
      var priorProbs= [1-priorReg, priorReg]
      var priorPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: prior_objects, ps: priorProbs}));
      return obj.shape == "circle" ? 1 : 0
      }})

      observe(priorPredictions, row.correct)
    }, conditionData)

    return {pref_int, pref_slope, condition: cndtn}

  }, all_conditions)

  var diff_speaker_params = _.filter(pref_params, 
        {condition: "different_speaker"})[0]
  var same_speaker_params = _.filter(pref_params, 
        {condition: "same_speaker"})[0]

  var predictions_by_age = map(function(age_bin){

    var speakerOptimality = so_int  + so_slope * age_bin

    var diff_priorReg = logistic(diff_speaker_params.pref_int +
            diff_speaker_params.pref_slope * age_bin)

    var same_priorReg = logistic(same_speaker_params.pref_int +
            same_speaker_params.pref_slope * age_bin)

    return extend(
      {
      diff_priorReg,
      same_priorReg,
      speakerOptimality,
      age_bin,

    })

  }, binned_age_range)

  return predictions_by_age
}
'
```

```{r}
params_by_age<- readRDS("../saves/params_by_age.rds")

# params_by_age<- webppl(
#   program_code = paste(rsaUtils, rsaModel, paramModelWebppl , sep='\n'),
#   data = pre_data,
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 200, verbose = T)
# )
# 
# saveRDS(params_by_age, file = "../saves/params_by_age.rds")

model_params_by_age <- params_by_age%>%
  select(value) %>%
  map_df(bind_rows)%>%
  group_by(age_bin)%>%
  summarise(so_map = estimate_mode(speakerOptimality),
            diff_priorReg_map = estimate_mode(diff_priorReg),
            same_priorReg_map = estimate_mode(same_priorReg))

```

## Basic RSA model predictions

```{r}
modelPredWebppl <- '

var levels = function(df, label){
  return _.uniq(_.map(df, label));
}

var ex3Data = dataFromR.ex3data
var modelParams = dataFromR.modelParams

    display(JSON.stringify(ex3Data))

var foreach = function(fn, lst) {
    var foreach_ = function(i) {
        if (i < lst.length) {
            fn(lst[i]);
            foreach_(i + 1);
        }
    };
    foreach_(0);
};

var logistic = function(x) {1 / (1 + Math.exp(-x))}

var predict_by_age = function(age_bin){

    //display(JSON.stringify(age_bin))

    var speakerOptimality = age_bin.so_map

    var diff_priorReg = age_bin.diff_priorReg_map

    var diff_priorProbs_congruent = [1-diff_priorReg, 1-diff_priorReg, diff_priorReg]

    var diff_priorProbs_incongruent = [diff_priorReg, diff_priorReg, 1-diff_priorReg]

    var rsaPredictions_diff_cong = expectation(pragmaticListener({label: "dax", point: 2 }, diff_priorProbs_congruent,   speakerOptimality))

    var rsaPredictions_diff_incong = expectation(pragmaticListener({label: "dax", point: 2 }, diff_priorProbs_incongruent, speakerOptimality))

    var same_priorReg =age_bin.same_priorReg_map

    var same_priorProbs_congruent = [1-same_priorReg, 1-same_priorReg, same_priorReg]

    var same_priorProbs_incongruent = [same_priorReg, same_priorReg, 1-same_priorReg]

    var rsaPredictions_same_cong = expectation(pragmaticListener({label: "dax", point: 2 }, same_priorProbs_congruent, speakerOptimality))

    var rsaPredictions_same_incong = expectation(pragmaticListener({label: "dax", point: 2 }, same_priorProbs_incongruent, speakerOptimality))

  return extend(
      {rsaPredictions_diff_cong,
      rsaPredictions_diff_incong,
      rsaPredictions_same_cong,
      rsaPredictions_same_incong,
      age_bin
    })

 }

var predictions_by_age = map(predict_by_age, modelParams)

predictions_by_age

'
```

```{r combined model predictions}

#model_pred <- readRDS("../saves/kids_model_predictions.rds")

model_pred<- webppl(
  program_code = paste(rsaUtils, rsaModel, modelPredWebppl , sep='\n'),
  data = list(ex3data = ex3_data, modelParams = model_params_by_age),
  data_var = "dataFromR",
  model_var = "predict_by_age",
  chains = 1,
  cores = 4,
  inference_opts = list(method = "MCMC", samples = 10, burn = 1, verbose = T)
)

#saveRDS(model_pred, file = "../saves/kids_model_predictions.rds")

```

## Noisy model predictions

```{r}
noisyModelPredWebppl <- '

var levels = function(df, label){
  return _.uniq(_.map(df, label));
}
var prefData = _.filter(dataFromR, {experiment: "preference_pretest"})
var infData = _.filter(dataFromR, {experiment: "informativeness_pretest"})

// make finer grained as needed
var binned_age_range = _.range(0, 2, 0.05)

var all_conditions = levels(prefData, "condition")

var foreach = function(fn, lst) {
    var foreach_ = function(i) {
        if (i < lst.length) {
            fn(lst[i]);
            foreach_(i + 1);
        }
    };
    foreach_(0);
};

var logistic = function(x) {1 / (1 + Math.exp(-x))}

var model  = function(){

  // global parameters

  var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var so_slope = uniformDrift({a: -2, b: 2, width: 0.4})
  var so_int = uniformDrift({a: -2, b: 2, width: 0.4})

  foreach(function(row){

    var age = row.age_num

    var speakerOptimality = so_int  + so_slope * (age - infData[0].minage)

    var inf_priorProbs = [.5, .5, .5]

    var rsaPredictions = pragmaticListener({label: "dax", point: 2 },
    inf_priorProbs, speakerOptimality)

    observe(rsaPredictions, row.correct)

  }, infData)

  var pref_params = map(function(cndtn){

    var conditionData = _.filter(prefData, {condition: cndtn})

    var pref_slope = uniformDrift({a: -2, b: 2, width: 0.4})
    var pref_int = uniformDrift({a: -2, b: 2, width: 0.4})

    foreach(function(row){

      var age = row.age_num

      var priorReg = logistic(pref_int + pref_slope * (age - row.minage))

      var priorProbs= [1-priorReg, priorReg]

      var priorPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: prior_objects, ps: priorProbs}));
      return obj.shape == "circle" ? 1 : 0
      }})

      observe(priorPredictions, row.correct)

    }, conditionData)

    return {pref_int, pref_slope, condition: cndtn}

  }, all_conditions)

  var diff_speaker_params = _.filter(pref_params, 
        {condition: "different_speaker"})[0]

  var same_speaker_params = _.filter(pref_params, 
        {condition: "same_speaker"})[0]

  var predictions_by_age = map(function(age_bin){

    var speakerOptimality = so_int  + so_slope * age_bin

    var diff_priorReg = logistic(diff_speaker_params.pref_int +
            diff_speaker_params.pref_slope * age_bin)

    var diff_priorProbs_congruent = [1-diff_priorReg, 1-diff_priorReg, diff_priorReg]

    var diff_priorProbs_incongruent = [diff_priorReg, diff_priorReg, 1-diff_priorReg]

    var rsaPredictions_diff_cong = pragmaticListener({label: "dax", point: 2 }, diff_priorProbs_congruent,   speakerOptimality)

    var noisyRsaPredictions_diff_cong = expectation(addNoise(rsaPredictions_diff_cong, noise))

    var rsaPredictions_diff_incong = pragmaticListener({label: "dax", point: 2 }, diff_priorProbs_incongruent, speakerOptimality)

 var noisyRsaPredictions_diff_incong = expectation(addNoise(rsaPredictions_diff_incong, noise))

    var same_priorReg = logistic(same_speaker_params.pref_int +
            same_speaker_params.pref_slope * age_bin)

    var same_priorProbs_congruent = [1-same_priorReg, 1-same_priorReg, same_priorReg]

    var same_priorProbs_incongruent = [same_priorReg, same_priorReg, 1-same_priorReg]

  var rsaPredictions_same_cong = pragmaticListener({label: "dax", point: 2 }, same_priorProbs_congruent, speakerOptimality)

 var noisyRsaPredictions_same_cong = expectation(addNoise(rsaPredictions_same_cong, noise))

   var rsaPredictions_same_incong = pragmaticListener({label: "dax", point: 2 }, same_priorProbs_incongruent, speakerOptimality)

 var noisyRsaPredictions_same_incong = expectation(addNoise(rsaPredictions_same_incong, noise))

    return [ 
    [ age_bin + "_diff_cong",  noisyRsaPredictions_diff_cong],
    [ age_bin + "_diff_incong",  noisyRsaPredictions_diff_incong],
    [ age_bin + "_same_cong",  noisyRsaPredictions_same_cong],
    [ age_bin + "_same_incong",  noisyRsaPredictions_same_incong],
]

  }, binned_age_range)

return _.fromPairs(predictions_by_age.concat([
["so_slope", so_slope],
["so_int", so_int],
["noise", noise]
]))

}
'
```
```{r noisy combined model predictions}

noisy_model_pred <- readRDS("../saves/kids_noisy_model_predictions.rds")

noisy_model_pred<- webppl(
  program_code = paste(rsaUtils, rsaModel, noisyModelPredWebppl , sep='\n'),
  data = pre_data,
  data_var = "dataFromR",
  model_var = "model",
  chains = 4,
  cores = 4,
  inference_opts = list(method = "MCMC", samples = 2000, burn = 200, verbose = T)
)

#saveRDS(noisy_model_pred, file = "../saves/kids_noisy_model_predictions.rds")

kids_noisy_model_pred <- noisy_model_pred%>%
  select(value) %>%
  map_df(bind_rows)%>%
  mutate(iteration = rep(1:8000,each = 40))%>%
  select(-diff_priorReg,-same_priorReg,-speakerOptimality,-so_slope,-so_int)%>%
  gather(condition, prop_informative, -iteration,-age_bin,-noise)%>%
  separate(condition, into = c("model", "Speaker", "Alignment"), sep="_")%>%
  mutate(Speaker = ifelse(Speaker == "diff", "different_speaker","same_speaker"),
         Alignment = ifelse(Alignment == "cong", "congruent","incongruent"))

```

## Prior only model

```{r prior only model}
priorOnlyWebppl <- '

var levels = function(df, label){
  return _.uniq(_.map(df, label));
}
var prefData = _.filter(dataFromR, {experiment: "preference_pretest"})

// make finer grained as needed
var binned_age_range = _.range(0, 2, 0.05)

var all_conditions = levels(prefData, "condition")

var foreach = function(fn, lst) {
    var foreach_ = function(i) {
        if (i < lst.length) {
            fn(lst[i]);
            foreach_(i + 1);
        }
    };
    foreach_(0);
};

var logistic = function(x) {1 / (1 + Math.exp(-x))}

var model  = function(){

  var pref_params = map(function(cndtn){

    var conditionData = _.filter(prefData, {condition: cndtn})

    var pref_slope = uniformDrift({a: -2, b: 2, width: 0.4})
    var pref_int = uniformDrift({a: -2, b: 2, width: 0.4})

    foreach(function(row){

      var age = row.age_num

      var priorReg = logistic(pref_int + pref_slope * (age - row.minage))

      var priorProbs= [1-priorReg, priorReg]

      var priorPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: prior_objects, ps: priorProbs}));
      return obj.shape == "circle" ? 1 : 0
      }})

      observe(priorPredictions, row.correct)

    }, conditionData)

    return {pref_int, pref_slope, condition: cndtn}

  }, all_conditions)

  var diff_speaker_params = _.filter(pref_params, 
        {condition: "different_speaker"})[0]

  var same_speaker_params = _.filter(pref_params, 
        {condition: "same_speaker"})[0]

  var predictions_by_age = map(function(age_bin){

    var diff_priorReg = logistic(diff_speaker_params.pref_int +
            diff_speaker_params.pref_slope * age_bin)

    var diff_priorProbs_congruent = [1-diff_priorReg, 1-diff_priorReg, diff_priorReg]

    var diff_priorProbs_incongruent = [diff_priorReg, diff_priorReg, 1-diff_priorReg]

    var same_priorReg = logistic(same_speaker_params.pref_int +
            same_speaker_params.pref_slope * age_bin)

    var same_priorProbs_congruent = [1-same_priorReg, 1-same_priorReg, same_priorReg]

    var same_priorProbs_incongruent = [same_priorReg, same_priorReg, 1-same_priorReg]

 var priorPredictions_diff_con = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: diff_priorProbs_congruent}));
      return obj.shape == "circle" ? 1 : 0
      }})

  var priorPredictions_diff_incon = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: diff_priorProbs_incongruent}));
      return obj.shape == "circle" ? 1 : 0
      }})

 var priorPredictions_same_con = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: same_priorProbs_congruent}));
      return obj.shape == "circle" ? 1 : 0
      }})

  var priorPredictions_same_incon = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: same_priorProbs_incongruent}));
      return obj.shape == "circle" ? 1 : 0
      }})

     var prior_diff_con = Math.exp(priorPredictions_diff_con.score(1))
var prior_diff_incon = Math.exp(priorPredictions_diff_incon.score(1))
     var prior_same_con = Math.exp(priorPredictions_same_con.score(1))
var prior_same_incon = Math.exp(priorPredictions_same_incon.score(1))

    return extend(
      {prior_diff_con,
prior_diff_incon,
prior_same_con,
prior_same_incon,
    age_bin})

  }, binned_age_range)

  return predictions_by_age 
}
'
```

```{r}

prior_only_predictions <- readRDS("../saves/prior_only_predictions.rds")

# prior_only_predictions<- webppl(
#   program_code = paste(rsaUtils, rsaModel, priorOnlyWebppl , sep='\n'),
#   data = pre_data,
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 3,
#   cores = 3,
#   inference_opts = list(method = "MCMC", samples = 1000, burn = 200, verbose = T)
# )

#saveRDS(prior_only_predictions, "../saves/prior_only_predictions.rds")

prior_only_model <- prior_only_predictions %>%
  select(value) %>%
  map_df(bind_rows)%>%
  mutate(iteration = rep(1:3000,each = 40))%>%
  gather(condition, prop_informative, -iteration,-age_bin)%>%
  separate(condition, into = c("model", "Speaker", "Alignment"), sep="_")%>%
  mutate(Speaker = ifelse(Speaker == "diff", "different_speaker","same_speaker"),
         Alignment = ifelse(Alignment == "con", "congruent","incongruent"))

```

# Sanity checks

### Model based speaker optimality by age

```{r plot so model pred, echo = F}

plot_model_pred_so <- kids_model_pred_so%>%
  group_by(age_bin)%>%
  summarise(mean = mean(speakerOptimality),
            lci = hdi_lower(speakerOptimality),
            uci = hdi_upper(speakerOptimality))

pred_so_plot <- ggplot(data = plot_model_pred_so) +
  geom_line(data = kids_model_pred_so,aes(age_bin, speakerOptimality, group = iteration), col = "grey", alpha = 0.3) +
  geom_ribbon(aes(x =age_bin, ymin = lci, ymax = uci), fill = "green", alpha = .2) +
  geom_line(aes(age_bin, mean), col = "black", size = 0.4, alpha = 1) + 
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0,0.5,1,1.5,2), labels = c(3.0,3.5,4.0,4.5,5)) +
  scale_y_continuous(limits=c(-4, 4), name="Speaker optimality")+
  ggtitle("Speaker optimality model")+
  #facet_wrap(~chain)+
  theme_few()+
  coord_fixed(ratio=1/4)

```

```{r, echo = F, fig.width= 8}
ggarrange(pred_so_plot, plot_inf)
```

Here we estimate the speaker optimality parameter that best captures the data for the different ages. The model based estimate nicely captures the developmental trend.

```{r plot prior model pred, echo = F}
plot_model_pred_prior <- kids_model_pred_prior%>%
  group_by(age_bin,condition)%>%
  summarise(mean = mean(prop_preferred),
            lci = hdi_lower(prop_preferred),
            uci = hdi_upper(prop_preferred))

pred_prior_plot <- ggplot(data = plot_model_pred_prior) +
  geom_line(data = kids_model_pred_prior,aes(age_bin, prop_preferred, group = paste(iteration,condition, sep= ""), col = condition), alpha = 0.05) +
  geom_ribbon(aes(x =age_bin, ymin = lci, ymax = uci, fill = condition), alpha = .5) +
  geom_line(aes(age_bin, mean, group = condition), col = "black", size = 0.4, alpha = 1) + 
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0,0.5,1,1.5,2), labels = c(3.0,3.5,4.0,4.5,5)) +
  scale_y_continuous(limits=c(-0.05,1.05), name="Proportion Expected Choice")+
  geom_hline(yintercept = 0.5, lty=2)+
  theme_few()+
  ggtitle("Preference model")+
  coord_fixed(ratio = 2)+
  scale_color_solarized()+
  scale_fill_solarized()

```

### Model based priors by age

```{r, echo= F, fig.width= 10}
ggarrange(pred_prior_plot, plot_pref, common.legend = T, legend = "right")
```

The model accurately captures the the difference between conditions as well as the developmental trend we see in the data.

# Visualizing model predictions

## RSA Model Predictions

These are the predictions from the RSA model which takes in the age specific speaker optimality parameter and the age specific priors. The priors are converted from 2-object world into 3-object world before going into the model. This corresponds to the way we handled the priors for adults.

```{r plot RSA model predictions, echo = F, fig.width= 10}

  plot_model_pred <- kids_model_pred %>%
  group_by(model,age_bin,Speaker,Alignment)%>%
  summarise(mean = mean(prop_informative),
            lci = hdi_lower(prop_informative),
            uci = hdi_upper(prop_informative))

pred_plot <- ggplot(plot_model_pred)+
  #geom_line(data = kids_model_pred,aes(age_bin, prop_informative, group = paste(iteration,Speaker, sep= ""), col = Speaker),  alpha = 0.05) + 
  #geom_pointrange(aes(y = mean, ymin = lci, ymax = uci))+
  geom_ribbon(aes(x =age_bin, ymin = lci, ymax = uci, fill = Speaker), alpha = .4) +
  geom_line(aes(age_bin, mean), col = "black", size = 0.4, alpha = 1) + 
  geom_hline(yintercept = 0.5, lty=2)+
  labs(y="Proportion Expected Choice")+
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0,0.5,1,1.5,2), labels = c(3.0,3.5,4.0,4.5,5)) +
  ylim(-0.05,1.05)+
  guides(col = F)+
  facet_grid(Speaker~Alignment)+
  theme_few()+
  scale_color_solarized()+
  scale_fill_solarized()
```

```{r, echo= F, fig.width= 10, fig.height = 8}
pred_plot
```

## Noisy RSA Model Predictions

```{r plot RSA model predictions, echo = F, fig.width= 10}

  plot_noisy_model_pred <- kids_noisy_model_pred %>%
  group_by(model,age_bin,Speaker,Alignment)%>%
  summarise(mean = mean(prop_informative),
            lci = hdi_lower(prop_informative),
            uci = hdi_upper(prop_informative))

pred_noisy_plot <- ggplot(plot_noisy_model_pred)+
  #geom_line(data = kids_model_pred,aes(age_bin, prop_informative, group = paste(iteration,Speaker, sep= ""), col = Speaker),  alpha = 0.05) + 
  #geom_pointrange(aes(y = mean, ymin = lci, ymax = uci))+
  geom_ribbon(aes(x =age_bin, ymin = lci, ymax = uci, fill = Speaker), alpha = .4) +
  geom_line(aes(age_bin, mean), col = "black", size = 0.4, alpha = 1) + 
  geom_hline(yintercept = 0.5, lty=2)+
  labs(y="Proportion Expected Choice")+
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0,0.5,1,1.5,2), labels = c(3.0,3.5,4.0,4.5,5)) +
  ylim(-0.05,1.05)+
  guides(col = F)+
  facet_grid(Speaker~Alignment)+
  theme_few()+
  scale_color_solarized()+
  scale_fill_solarized()
```

```{r, echo= F, fig.width= 10, fig.height = 8}
pred_noisy_plot
```

## No Prior Model Predictions

```{r, include = F}

no_prior_pred <- bind_rows(
  kids_model_pred_so%>%mutate(model = "no prior",
                              Speaker = "same_speaker",
                              Alignment = "congruent"),
    kids_model_pred_so%>%mutate(model = "no prior",
                              Speaker = "different_speaker",
                              Alignment = "congruent"),
    kids_model_pred_so%>%mutate(model = "no prior",
                              Speaker = "same_speaker",
                              Alignment = "incongruent"),
    kids_model_pred_so%>%mutate(model = "no prior",
                              Speaker = "different_speaker",
                              Alignment = "incongruent")
  ) %>%
  mutate(prop_informative = plogis(so_int+so_slope*age_bin))

no_prior_plot <- no_prior_pred %>%
  group_by(model,age_bin,Speaker,Alignment)%>%
  summarise(mean = mean(prop_informative),
            lci = hdi_lower(prop_informative),
            uci = hdi_upper(prop_informative))

plot_no_prior <- ggplot(no_prior_plot)+
  geom_line(data = no_prior_pred,aes(age_bin, prop_informative, group = paste(iteration,Speaker, sep= ""), col = Speaker),  alpha = 0.05) + 
  #geom_pointrange(aes(y = mean, ymin = lci, ymax = uci))+
  geom_ribbon(aes(x =age_bin, ymin = lci, ymax = uci, fill = Speaker), alpha = .4) +
  geom_line(aes(age_bin, mean), col = "black", size = 0.4, alpha = 1) + 
  geom_hline(yintercept = 0.5, lty=2)+
  labs(y="Proportion Expected Choice")+
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0,0.5,1,1.5,2), labels = c(3.0,3.5,4.0,4.5,5)) +
  ylim(-0.05,1.05)+
  guides(col = F)+
  facet_grid(Speaker~Alignment)+
  theme_few()+
  scale_color_solarized()+
  scale_fill_solarized()
```

```{r, echo = F, fig.width= 10, fig.height = 8}
plot_no_prior
```

## Prior Only Model

```{r plot prior only model predictions, include= F}

plot_prior_only_pred <- prior_only_model %>%
  group_by(model,age_bin,Speaker,Alignment)%>%
  summarise(mean = mean(prop_informative),
            lci = hdi_lower(prop_informative),
            uci = hdi_upper(prop_informative))

prior_only_plot <- ggplot(plot_prior_only_pred)+
  geom_line(data = prior_only_model,aes(age_bin, prop_informative, group = paste(iteration,Speaker, sep= ""), col = Speaker),  alpha = 0.05) + 
  #geom_pointrange(aes(y = mean, ymin = lci, ymax = uci))+
  geom_ribbon(aes(x =age_bin, ymin = lci, ymax = uci, fill = Speaker), alpha = .4) +
  geom_line(aes(age_bin, mean), col = "black", size = 0.4, alpha = 1) + 
  geom_hline(yintercept = 0.5, lty=2)+
  labs(y="Proportion Expected Choice")+
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0,0.5,1,1.5,2), labels = c(3.0,3.5,4.0,4.5,5)) +
  ylim(-0.05,1.05)+
  facet_grid(Speaker~Alignment)+
  theme_few()+
  scale_color_solarized()+
  scale_fill_solarized()

```

```{r, echo = F, fig.width= 10, fig.height = 8}
prior_only_plot
```

## Comparing model predictions

```{r}
model_comp_mean_ci <- bind_rows(
plot_model_pred%>%ungroup()%>%mutate(model = "RSA"), 
plot_prior_only_pred%>%ungroup()%>%mutate(model = "Prior Only"),
no_prior_plot%>%ungroup()%>%mutate(model = "No Prior")
)

#saveRDS(model_comp_mean_ci, "../saves/kids_model_comp_mean_ci.rds")

model_comp <- bind_rows(
kids_model_pred %>%mutate(model = "RSA"),
prior_only_model%>%mutate(model = "Prior Only"),
no_prior_pred%>%select(-so_slope,-so_int,-speakerOptimality)%>%mutate(model = "No Prior")
  )

# saveRDS(model_comp, "../saves/kids_model_comp.rds")

```

```{r, include = F}
pred_plot <- ggplot(model_comp_mean_ci)+
  geom_line(data = model_comp,aes(age_bin, prop_informative, group = paste(iteration,Speaker,model, sep= ""), col = model),  alpha = 0.01) + 
  geom_ribbon(aes(x =age_bin, ymin = lci, ymax = uci, fill = model), alpha = .4) +
  geom_line(aes(age_bin, mean, col = model), size = 1, alpha = 1) + 
  geom_hline(yintercept = 0.5, lty=2)+
  labs(y="proportion more informative")+
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0,0.5,1,1.5,2), labels = c(3.0,3.5,4.0,4.5,5)) +
  ylim(-0.05,1.05)+
  facet_grid(Speaker~Alignment)+
  theme_few()+
  scale_color_viridis_d()+
  scale_fill_viridis_d()
```

```{r, echo = F, fig.width= 10, fig.height = 8}
pred_plot
```

```{r, include = F}
#ggsave("../graphs/no_prior_prior_only_model_predictions.pdf", width = 10, height = 8, scale = 1)
```

# Model vs data

```{r import data, include = F}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(jsonlite)
library(readxl)

## importing data
ex3_data <- read_csv(file="../data/kids_ex3_data.csv")

# files <- dir("~/Work/MCC/git-mcc/kids_EX3/pref/")
# 
# raw_data <- data_frame()
# for (f in files) {
#   jf <- paste("~/Work/MCC/git-mcc/kids_EX3/pref/",f,sep="")
#   jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
#   date <- str_sub(jf,42,str_length(jf)-11)
#   id <- as_data_frame(jd$data$data) %>% mutate(test_date = date)
#   raw_data <- bind_rows(raw_data, id)
# }
# 
# raw_data <- raw_data %>%
#   mutate(subid = ifelse(subid == "181128_3_ex3p....","181128_3_ex3p" , subid),
#          subid = ifelse(subid == "181128_1_ex3p.","181128_1_ex3p" , subid),
#          subid = ifelse(subid == " 181218_10_ex3p","181218_10_ex3p" , subid))
# 
# raw_data%>%
#   #group_by(paste(subid,test_date))%>%
#   group_by(subid)%>%
#   summarise(n= n())
#   
# unique(raw_data$subid)
# 
# log <- read_excel("../../../MCC-subject_log.xlsx", 1)%>%
#   filter(Condition == "ex3p")%>%
#   select(subid,experimenter,keep_drop,sex,dob)
# 
# #unique(log$subid)
# sum(log$keep_drop == "drop")
# 
# d <- left_join(raw_data,log, by = "subid")%>%
#   filter(keep_drop == "keep")%>%
#   mutate(dob = as.numeric(dob),
#          dob = as.Date(dob,origin = "1899-12-30"),
#          age_num = lubridate::time_length(difftime(test_date,dob), "years"),
#          check_age = ifelse(substr(age_num,1,1) == subage, T, F),
#          Speaker = ifelse(speaker == "same", "same_speaker", "different_speaker"),
#          Alignment = alignment)%>%
#   select(-speaker,-alignment)
# 
# 
# #write_csv(d, "../data/kids_ex3_data.csv")

```

## Plot comparing model and data by continous age
```{r comparison plot by age, echo= FALSE}

p_data <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  mutate(age_bin = age_num - 3)%>%
  group_by(Speaker, Alignment ,age_bin, subid)

ggplot(model_comp_mean_ci)+
  geom_jitter(data = p_data, aes(x = age_bin, y = correct_inf), width = .025,height = .025, alpha = .2)+
  geom_ribbon(aes(x =age_bin, ymin = lci, ymax = uci, fill = model), alpha = .4) +
  geom_line(aes(age_bin, mean, col = model), size = 1, alpha = 1, linetype = 2) + 
  geom_smooth(data = p_data, aes(x = age_bin, y = correct_inf), col = "black", size = 0.6, method = "glm", method.args = list(family = "binomial"),  se = T,  alpha = .5, size = .4)+
  #geom_hline(yintercept = 0.5, lty=2)+
  labs(y="proportion more informative")+
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0,0.5,1,1.5,2), labels = c(3.0,3.5,4.0,4.5,5)) +
  ylim(-0.05,1.05)+
  facet_grid(Speaker~Alignment)+
  theme_few()+
  scale_color_solarized()+
  scale_fill_solarized()

```

## Correlation plot
```{r}

data <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by(Speaker, Alignment ,subage) %>%
  summarize(k = sum(correct_inf), n = n())%>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         Data_ci_lower  = qbeta(.025, a, b),
         Data_ci_upper = qbeta(.975, a, b),
         Data = (a-1)/(a+b-2))%>%
  select(-a,-b,-n,-k)

model <- model_comp %>%
  mutate(subage = ifelse(age_bin<1,"3","4"))%>%
  group_by(model,Speaker,Alignment,subage)%>%
  summarise(Model = mean(prop_informative),
            Model_ci_lower = hdi_lower(prop_informative),
            Model_ci_upper = hdi_upper(prop_informative))

cor_plot <- model %>%
  ungroup()%>%
  mutate(Data = rep(data$Data,3),
         Data_ci_lower = rep(data$Data_ci_lower,3),
         Data_ci_upper  =rep(data$Data_ci_upper,3))

ggplot(data = cor_plot,aes(x = Model, y = Data, col = subage)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_point(size = 3)+
  geom_errorbar(aes(ymin = Data_ci_lower, ymax = Data_ci_upper),width = 0,size = .5)+
  geom_errorbarh(aes(xmin = Model_ci_lower, xmax = Model_ci_upper), height = 0,size = .5)+
  coord_fixed()+
  stat_cor(method = "pearson", label.x = 0.1, label.y = 0.9, aes(x = Model, y = Data), inherit.aes = F)+
  facet_grid(~model)+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  theme_few(base_size = 10) + 
  scale_colour_solarized()

cor.test(cor_plot$Model,cor_plot$Data)
```