### 2019-02-24 stats/kids/kids_inf_biling.Rmd
kids model prediction with replication data

---
title: "MCC Informativeness Kids - Additional data"
author: "Manuel Bohn"
output: html_document
---

```{r import data, include = F}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(jsonlite)
library(readxl)

## importing data
files <- dir("~/Work/MCC/git-mcc/kids_info_data_2/")

raw_data_2 <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_data_2/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,47,str_length(jf)-11)
  id <- as_data_frame(jd$data$data) %>% mutate(test_date = date)
  raw_data_2 <- bind_rows(raw_data_2, id)
}

d_raw_2 <-raw_data_2 %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  mutate(test_date = as.Date(test_date),
         experiment = "mcc_kids_inf_2_2",
         subid= ifelse(subid=="193101_1_inf2","190131_1_inf2",subid))%>%
  select(subid,subage,trial,correct,test_date)

files <- dir("~/Work/MCC/git-mcc/kids_info_data/")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_data/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,45,str_length(jf)-11)
  id <- as_data_frame(jd$data$data) %>% mutate(test_date = date)
  raw_data <- bind_rows(raw_data, id)
}

d_raw <-raw_data %>%
    filter(trial != "filler1",
         trial != "filler2")%>%
  mutate(test_date= ifelse(test_date == "18-10-10", paste(20,test_date,sep = ""),test_date),
         subid= ifelse(subid == "101018_8_inf", "181010_8_inf",subid),
         subid= ifelse(subid == "181031_3", "181031_3_inf",subid),
         test_date= ifelse(test_date == "18-XX-XX", "2018-10-10",test_date),
         test_date= ifelse(test_date == "18-09-12", "2018-09-12",test_date),
         experiment = "mcc_kids_inf_2")%>%
  select(subid,subage,trial,correct,test_date)%>%
  mutate(test_date = as.Date(test_date))

files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/trees")

raw_data_trees <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/trees/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,50,str_length(jf)-11)
  id <- as_data_frame(jd$data$data)%>% mutate(test_date = date)
  raw_data_trees <- bind_rows(raw_data_trees, id)
}

d_trees <- raw_data_trees %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "filler1",
         trial != "filler2",
         subage != "Test")%>%
  mutate(test_date = as.Date(test_date))%>%
  select(subid,subage,trial,correct,test_date)

files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/robots_point_4_map")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/robots_point_4_map/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,63,str_length(jf)-11)
  id <- as_data_frame(jd$data$data)%>% mutate(test_date = date)
  raw_data <- bind_rows(raw_data, id)
}

d_robots <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "filler1",
         trial != "filler2",
         subage != "Test")%>%
  mutate(test_date = as.Date(test_date))%>%
  select(subid,subage,trial,correct,test_date)

files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/trees_left_right")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/trees_left_right/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,61,str_length(jf)-11)
  id <- as_data_frame(jd$data$data)%>% mutate(test_date = date)
  raw_data <- bind_rows(raw_data, id)
}

d_trees_left_right <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "filler1",
         trial != "filler2",
         subage != "Test")%>%
  mutate(test_date = as.Date(test_date))%>%
  select(subid,subage,trial,correct,test_date)

d_sum <- bind_rows(
  d_raw_2,
  d_raw,
  d_trees,
  d_robots,
  d_trees_left_right)

log <- read_excel("../../../MCC-subject_log.xlsx", 1)%>%
  filter(Condition == "inf" | Condition == "inf2")%>%
  select(subid,experimenter,keep_drop,sex,dob, English, parent1_ed,parent2_ed)%>%
  mutate(language_status = ifelse(English == "100" | English == "90-95", "mono_ling", "bi_ling"))

d <- left_join(d_sum,log, by = "subid")%>%
  mutate(dob = as.numeric(dob),
         dob = as.Date(dob,origin = "1899-12-30"),
         age_num = lubridate::time_length(difftime(test_date,dob), "years"),
         check_age = ifelse(substr(age_num,1,1) == subage, T, F))%>%
  filter(language_status != "NA", 
         English != "NA")

write_csv(d, "../data/kids_inf_data_megan.csv")

```

# Sample size per age bin

```{r kids per age group,echo = F }

d %>%
  group_by(subage, sex)%>%
  summarise(n = length(unique(subid)))

```

# Plot test trials by age bin

```{r plot by age, echo= FALSE}

p1 <- d %>%
  group_by(sex,subage, subid) %>%
  summarise(correct = mean(correct))

p2 <- p1 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p1, aes(x = sex, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = sex, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized()
```

# Plot with age as continuous variable

```{r plot by numerical age, echo = F}
p5 <- d %>%
  group_by(language_status,age_num, subid) %>%
  summarise(correct = mean(correct))

ggplot(data = p5, aes(x = age_num, y = correct,col = language_status)) +
  geom_smooth(method = "lm", se = T)+
  geom_jitter( alpha = .5, height = 0.04)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  #theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized()
```

# Performance by trial and age bin

```{r plot trials, echo = F}
ms <- d %>%
  filter(trial != "filler1",
         trial != "filler2") %>%
  group_by(subage, trial, subid) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(ms, 
       aes(x = trial, y = mean, col = subage)) +
  geom_line(aes(group= subage)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  theme_few() + 
  scale_colour_solarized()
```

## T-tests against chance per age bin
```{r, echo = F}
## t tests against chance
d %>%
  filter(trial != "filler1",
         trial != "filler2") %>%
  group_by(subage, subid) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(subage) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(subage,df,t_value,p_value) %>%
  knitr::kable(digits = 3)

```

## GLMM

```{r, include=F}
## model
library(lme4)

lm_data <- d%>%
  filter(trial != "filler1",
         trial != "filler2") %>%
  mutate(age_num = scale(age_num, center = TRUE, scale = TRUE))

# maximal converging model
lm <- glmer(correct ~ age_num + (1 | subid), 
              data = lm_data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

```

```{r, echo = F}
summary(lm)
```

# Effect of target position

Target was either the closer (inner) or further away (outer) object.

```{r, echo = F}
p5 <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by( subage, targetPosition, subid) %>%
  summarise(correct = mean(correct))

p6 <- p5 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p5, aes(x = targetPosition, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p6, aes(x = targetPosition, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

# Target switch

In half of the trials, the target switched sides from exposure to test.

```{r, echo = F}
p7 <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by( subage, targetSwitch, subid) %>%
  summarise(correct = mean(correct))

p8 <- p7 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p7, aes(x = targetSwitch, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p8, aes(x = targetSwitch, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

# Joining round 1 and round 2

```{r}
files <- dir("~/Work/MCC/git-mcc/kids_info_data/")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_data/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,45,str_length(jf)-11)
  id <- as_data_frame(jd$data$data) %>% mutate(test_date = date)
  raw_data <- bind_rows(raw_data, id)
}

d_raw <-raw_data %>%
  mutate(test_date= ifelse(test_date == "18-10-10", paste(20,test_date,sep = ""),test_date),
         subid= ifelse(subid == "101018_8_inf", "181010_8_inf",subid),
         subid= ifelse(subid == "181031_3", "181031_3_inf",subid),
         test_date= ifelse(test_date == "18-XX-XX", "2018-10-10",test_date),
         test_date= ifelse(test_date == "18-09-12", "2018-09-12",test_date),
         experiment = "mcc_kids_inf_2")

log <- read_excel("../../../MCC-subject_log.xlsx", 1)%>%
  filter(Condition == "inf")%>%
  select(subid,experimenter,keep_drop,sex,dob)

d2 <- left_join(d_raw,log, by = "subid")%>%
  filter(keep_drop == "keep")%>%
  mutate(dob = as.numeric(dob),
         dob = as.Date(dob,origin = "1899-12-30"),
         age_num = lubridate::time_length(difftime(test_date,dob), "years"),
         check_age = ifelse(substr(age_num,1,1) == subage, T, F))

d_sum <- bind_rows(d%>%select(-test_date),d2%>%select(-test_date)%>%filter(subage!="2"))%>%
    filter(trial != "filler1",
         trial != "filler2")

```

```{r}
d_sum %>%
  group_by(subage)%>%
  summarise(n = length(unique(subid)))%>%
  knitr::kable(digits = 2)
```

```{r plot by age, echo= FALSE}

p_sum1 <- d_sum %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by(control,subage, subid) %>%
  summarise(correct = mean(correct))

p_sum2 <- p_sum1 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p_sum1, aes(x = control, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p_sum2, aes(x = control, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

# Plot with age as continuous variable

```{r plot by numerical age, echo = F}
p_sum3 <- d_sum %>%
  filter(trial != "filler1",
         trial != "filler2") %>%
  group_by(age_num, subid) %>%
  summarise(correct = mean(correct))

ggplot(data = p_sum3, aes(x = age_num, y = correct)) +
  geom_smooth(method = "lm", se = T)+
  geom_point( alpha = .5, width = 0.02)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  #theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

```{r, include=F}
## model
library(lme4)

lm_data_sum <- d_sum%>%
  mutate(age_num = scale(age_num, center = TRUE, scale = TRUE))

# maximal converging model
lm_sum <- glmer(correct ~ age_num + (1 | subid), 
              data = lm_data_sum, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

```

```{r, echo = F}
summary(lm_sum)
```