### 2019-04-26 papers/full_paper/models.Rmd
first draft paper si

---
title             : "Predicting information integration in pragmatic word learning across development"
shorttitle        : "Supplementary Information"

author: 
  - name          : "Manuel Bohn"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Psychology, 450 Serra Mall, Stanford, CA 94305"
    email         : "bohn@stanford.edu"
  - name          : "Michael Henry Tessler"
    affiliation   : "3"
  - name          : "Megan Merrick"
    affiliation   : "1"
  - name          : "Michael C. Frank"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Stanford University"
  - id            : "2"
    institution   : "Leipzig Research Center for Early Child Development, Leipzig University"
  - id            : "3"
    institution   : "Department of Brain and Cognitive Sciences, Massachusetts Institute of Technology"

bibliography      : ["library.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, include = T)

library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
library(coda)
library(matrixStats)
library(ggpubr)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}

```

# Model code

## Model utilities

```{r rsaUtils, include = T}
rsaUtils <- '
// three objects: triangle_1 on table 1, triangle_2 and circle_1 on table 2
var all_objects = [
{ shape: "triangle", id:1, location: 1},  
{ shape: "triangle", id:2, location: 2},
{ shape: "circle", id:1, location: 2}
]

// listener prior is over types, not tokens (we use uniformDraw on prior_objects)
var prior_objects = [
{ shape: "triangle", id:1, location: 1},  
{ shape: "circle", id:1, location: 2}
]

var labels = ["dax","wug"]

// two possible lexica, corresponding to meaning of words "dax" and "wug"
var lexicon1 = function(utterance, obj){
utterance.label == "dax" ? obj.shape == "triangle" :
utterance.label == "wug" ? obj.shape == "circle" : 
true
}

var lexicon2 = function(utterance, obj){
utterance.label == "dax" ? obj.shape == "circle" :
utterance.label == "wug" ? obj.shape == "triangle" : 
true
}

var lexiconObjects = {
"dax = triangle": {
triangle: "dax", circle: "wug"
},
"dax = circle": {
triangle: "wug", circle: "dax"
},
}

var lexiconObject = {
"dax = triangle": lexicon1,
"dax = circle" : lexicon2
}

var point = function(utterance, obj){
return obj.location == utterance.point
}

var utterancePrior = function(obj, lexiconName){
  var locationsWithShape = _.map(_.filter(all_objects, {shape: obj.shape}), "location")
  var point = uniformDraw(locationsWithShape)
  var label = lexiconObjects[lexiconName][obj.shape]
  return {label: label, point: point}
}

var LexiconPrior = Categorical({vs: ["dax = triangle","dax = circle" ], ps: [1, 1]})

var addNoise = function(dist, noiseParam){
   Infer({model: function(){ 
      return flip(noiseParam) ? uniformDraw([0, 1]) : sample(dist)
    }
   })
}

var foreach = function(fn, lst) {
    var foreach_ = function(i) {
        if (i < lst.length) {
            fn(lst[i]);
            foreach_(i + 1);
        }
    };
    foreach_(0);
};

var logistic = function(x) {1 / (1 + Math.exp(-x))}

var levels = function(df, label){
  return _.uniq(_.map(df, label));
}

'
```

## RSA model code

```{r rsa model, include = T}
rsaModel <- '
var literalListener = cache(function(utterance, priorProbs){
  Infer({method: "enumerate", model: function(){
    var lexiconName = sample(LexiconPrior); 
    var lexicon = lexiconObject[lexiconName];
    var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
    if ("label" in utterance) {
      var truthValue = lexicon(utterance, obj);
      condition(truthValue)
    }
    if (utterance.point) {
     var truthValuePoint = point(utterance, obj);
     condition(truthValuePoint)
    }
    return obj.shape 
  }})
})

var speaker = cache(function(obj, lexiconName, priorProbs, speakerOptimality){
  Infer({method: "enumerate", model: function(){
    var utterance = utterancePrior(obj, lexiconName);
    var L0 = literalListener(utterance, priorProbs);
    factor(speakerOptimality * L0.score(obj.shape))
    return utterance
  }})
})

var pragmaticListener = cache(function(utterance, priorProbs, speakerOptimality){
  Infer({method: "enumerate", model: function(){
    var lexiconName = sample(LexiconPrior);
    var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
    var S1 = speaker(obj, lexiconName, priorProbs, speakerOptimality);
    observe(S1, utterance)
    return obj.shape == "circle" ? 1 : 0
  }})
})

var addNoise = function(dist, noiseParam){
   Infer({model: function(){ 
      return flip(noiseParam) ? uniformDraw([0, 1]) : sample(dist)
    }
   })
}
'
```

# Adults

## Experiment 3

### Model parameters and data

```{r ex3 data adults}
# load data from experiment 3
adult_ex3_data <-read_csv(file="../../stats/data/adult_ex3.csv") 
```

```{r ex 1 and 2 data adults}
# adults exp 1 
adult_ex1_data <- read_csv(file="../../stats/data/adult_ex1.csv")
# load data from prior manipulation experiments to set priors
adult_ex2_novelty_data <- read_csv(file="../../stats/data/adult_ex2_novelty.csv")
adult_ex2_preference_data <- read_csv(file="../../stats/data/adult_ex2_preference.csv")
```

#### Inferring speaker optimality parameter and setting priors

```{r model parameter adults, include = T}
# Cauchy prior on the speaker optimality parameter around 1, based on previous work (e.g. Frank & Goodman, 2014) 
speakOptParam <- '
var data = dataFromR.data

var model  = function(){

  var speakerOptimality = sample(Cauchy({location: 1, scale: 2}))

  foreach(function(row){

    var priorProbs = [.5, .5, .5]

    var rsaPredictions = pragmaticListener({label: "dax", point: 2 },
      priorProbs, speakerOptimality)

    observe(rsaPredictions, row.correct)
  }, data)

  return speakerOptimality

}
'
```

```{r inferring speaker optimality parameter adults}
# adult_model_param<- webppl(
#   program_code = paste(rsaUtils, rsaModel, speakOptParam , sep='\n'),
#   data = list(data = adult_ex1_data %>% filter(condition == "test")),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# )%>%
#   saveRDS("../../stats/saves/adult_model_param.rds")

adult_model_param <- readRDS("../../stats/saves/adult_model_param.rds")%>%
  select(value) %>%
  summarise(so_map = estimate_mode(value),
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

```

```{r visualize speaker optimality parameter adults, echo = T}
# visualize speaker optimality parameter
plot_adult_model_speak_opt <- readRDS("../../stats/saves/adult_model_param.rds")%>%
  select(value)

ggplot(plot_adult_model_speak_opt, aes(value))+
  geom_density()+
  xlab("Speaker optimality")+
  theme_few()
```

```{r model setup ex 3 adults}
# create datafiles with model parameters
model_params_ex3 <- expand.grid(
  common_ground_manipulation = c("preference", "novelty"),
  speaker = c("same_speaker", "different_speaker"),
  speak_opt = adult_model_param$so_map
  )

prior_pref_summary <- bind_rows(
  adult_ex2_preference_data,
  adult_ex2_novelty_data) %>%
  group_by(condition, experiment) %>%
  summarise(prior = mean(correct)) %>%
  mutate(experiment = gsub("adult_ex2_", "", experiment)) %>%
  rename(common_ground_manipulation = experiment,
         speaker = condition)

model_params_ex3 <- left_join( 
  model_params_ex3, 
  prior_pref_summary, by = c("common_ground_manipulation", "speaker")
) %>%
  rename(congruent = prior) %>%
  mutate(incongruent = 1 - congruent) %>%
  gather(alignment, prior, congruent, incongruent)

```

### Model predictions

#### Pragmatic model

##### Parameter free

```{r model rsa ex3 adults, include=T}
predictions.rsa.ex3 <-'

var allData = dataFromR.data

var output = map(function(row){

    var priorProbLst = [1 - row.prior, 1 - row.prior, row.prior]

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, priorProbLst, row.speak_opt)

    return extend([row.common_ground_manipulation + "/" + row.speaker + "/" + row.alignment, Math.exp(modelPredictions.score(1))])

}, allData)

output

'
```

```{r predictions pragmatic model ex 3 adults}
ex3_pragm_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex3 , sep='\n'),
  data =list(data = model_params_ex3),
  data_var = "dataFromR"
)

# model predictions
ex3_pragm_model_predictions <- ex3_pragm_model %>%
  separate(`0`, into = c("common_ground_manipulation", "speaker", "alignment"), sep="/")%>%
  mutate(value = `1`,
         model="pragmatic")%>%
  select(-`1`)

```

##### Noise parameter

```{r model rsa noise ex3, include = T}
predictions.rsa.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

var model = function(){

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [1 - conditionInfo.prior, 1 - conditionInfo.prior, conditionInfo.prior]

      var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors, conditionInfo.speak_opt)

   var noisyModelPredictions = addNoise(modelPredictions, noise)

 map(function(d){ 
      observe(noisyModelPredictions, d.correct_inf)
    }, conditionSpecificData)

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allParams)

  return extend(_.fromPairs(conditionOutput), {noise: noise})
}
'
```

```{r predictions pragmatic model ex 3 noise, include = FALSE}
# ex3_pragm_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_pragm_model_noise.rds")

# model predictions
ex3_pragm_model_pred_noise <- readRDS("../../stats/saves/ex3_pragm_model_noise.rds") %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation","speaker", "alignment"), sep="/")%>%
  mutate(model="pragmatic_noise")%>%
  group_by(model,common_ground_manipulation,speaker,alignment)%>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

## noise parameter
ex3_pragm_model_noise_param <- readRDS("../../stats/saves/ex3_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))
```

#### Prior only model

##### Noise parameter

```{r model prior only noise ex3, include = FALSE}
predictions.prior.only.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

var model = function(){

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [1 - conditionInfo.prior, 1 - conditionInfo.prior, conditionInfo.prior]

  var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: conditionSpecificPriors}));
      return obj.shape == "circle" ? 1 : 0
  }})

   var noisyModelPredictions = addNoise(modelPredictions, noise)

 map(function(d){ 
      observe(noisyModelPredictions, d.correct_inf)
    }, conditionSpecificData)

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allParams)

  return extend(_.fromPairs(conditionOutput), {noise: noise})
}
'
```

```{r predictions prior only model ex 3 noise, include = FALSE}
# ex3_prior_only_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.prior.only.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_prior_only_model_noise.rds")

# model predictions
ex3_prior_only_model_pred_noise <- readRDS("../../stats/saves/ex3_prior_only_model_noise.rds") %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation","speaker", "alignment"), sep="/")%>%
  mutate(model="prior_only_noise")%>%
  group_by(model,common_ground_manipulation,speaker,alignment)%>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

## noise parameter
ex3_prior_only_model_noise_param <- readRDS("../../stats/saves/ex3_prior_only_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))
```

#### Flat prior model

##### Noise parameter

```{r model flat prior noise ex3, include = T}
predictions.flat.prior.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

var model = function(){

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [.5, .5, .5]

      var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors, conditionInfo.speak_opt)

   var noisyModelPredictions = addNoise(modelPredictions, noise)

 map(function(d){ 
      observe(noisyModelPredictions, d.correct_inf)
    }, conditionSpecificData)

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allParams)

  return extend(_.fromPairs(conditionOutput), {noise: noise})
}
'
```

```{r predictions flat prior model ex 3 noise, include = FALSE}
# ex3_flat_prior_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.flat.prior.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_flat_prior_model_noise.rds")

# model predictions
ex3_flat_prior_model_pred_noise <- readRDS("../../stats/saves/ex3_flat_prior_model_noise.rds") %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation","speaker", "alignment"), sep="/")%>%
  mutate(model="flat_prior_noise")%>%
  group_by(model,common_ground_manipulation,speaker,alignment)%>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

## noise parameter
ex3_flat_prior_model_noise_param <- readRDS("../../stats/saves/ex3_flat_prior_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))
```

#### Compare model predictions to data

##### Noise parameters

```{r, include = T}
ex3_noise_parameters <- bind_rows(
  readRDS("../../stats/saves/ex3_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  mutate(model = "pragmatic"),
  readRDS("../../stats/saves/ex3_prior_only_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  mutate(model = "prior_only"),
  readRDS("../../stats/saves/ex3_flat_prior_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  mutate(model = "flat_prior")
  )

ggplot(ex3_noise_parameters, aes(x = value, col = model, fill = model))+
  geom_density(alpha = .5)+
  theme_few()+
  xlim(0,1)+
  scale_color_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#b58900","#6c71c4","#859900"))+
  scale_fill_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#b58900","#6c71c4","#859900"))
```

```{r}
# combine model predictions with data

# summarize data 
adult_ex3_data_summary <- adult_ex3_data %>%
  mutate(model = "data") %>%
  group_by(model,common_ground_manipulation,speaker,alignment)%>%
  summarize(k = sum(correct_inf), n = n())%>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         ci_lower  = qbeta(.025, a, b),
         ci_upper = qbeta(.975, a, b),
         mean = (a-1)/(a+b-2))%>%
  select(-a,-b,-n,-k)

# summarize model predictions
ex_3_model_pred_param_free <- bind_rows(
  ex3_pragm_model_predictions
)%>%
  group_by(model,common_ground_manipulation,speaker,alignment)%>%
  summarise(mean = mean(value))%>%
  mutate(ci_lower = mean,
         ci_upper = mean)

ex_3_model_pred <- 
  bind_rows(
    ex_3_model_pred_param_free,
    ex3_pragm_model_pred_noise,
    ex3_prior_only_model_pred_noise,
    ex3_flat_prior_model_pred_noise)

# combine model predictions and data

ex_3_pred <- bind_rows(
  ex_3_model_pred,
  adult_ex3_data_summary
)

#saveRDS(ex_3_pred, "../../stats/saves/adult_ex3_model_predictions_data.rds")

```

##### Plot model predictions and data

```{r plot ex3 model predictions and data adults, include = T}
ggplot(data = ex_3_pred, aes(x = alignment, y = mean, col = model)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), size = .4,  position = position_dodge(width = .5), pch = 5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion more informative")+
  facet_grid (~ common_ground_manipulation + speaker) +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_color_manual(name="Model",
                     breaks=c("data","pragmatic","pragmatic_noise","flat_prior_noise","prior_only_noise"),
                     values= c("black","#859900","#00afaf","#6c71c4","#b58900"))+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))
```

##### Correlation plot model predictions and data

```{r correlation plot ex3 adults}

ex_3_cor_plot <- ex_3_pred %>%
  ungroup()%>%
  filter(model != "data") %>%
  left_join(., ex_3_pred %>%
    ungroup()%>%
    filter(model == "data") %>%
    rename(data_mean = mean, data_ci_lower = ci_lower, data_ci_upper = ci_upper) %>%
    select(-model)
  )

ggplot(data = ex_3_cor_plot,aes(x = mean, y = data_mean, col = "#BB5566")) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 0.7, size = 0.5)+
  geom_point(size = 2)+
  geom_errorbar(aes(ymin = data_ci_lower, ymax = data_ci_upper),width = 0,size = .7)+
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0,size = .7)+
  coord_fixed()+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  ylab("Data")+
  guides(col = F)+
  facet_grid ( ~ model) +
  stat_cor(method = "pearson", label.x = 0.01, label.y = 0.99, aes(x = mean, y = data_mean), inherit.aes = F, size = 3)+
  theme_few()

```

### Model comparison

#### Marginal likelihoods

##### Pragmatic model

###### Parameter free

```{r log likes ex3 adults}
# calculating model log-likelihoods
# log_like_pragmatic_ex3 <- adult_ex3_data %>%
#   left_join(ex3_pragm_model_predictions)%>%
#   mutate(like = ifelse(correct_inf == 1, value, 1 - value),
#          log_like = log(like))%>%
#   group_by(model)%>%
#   summarise(logP = sum(log_like))%>%
#   saveRDS("../../stats/saves/ex3_pragm_model_loglike.rds")

log_like_pragmatic_ex3 <- readRDS("../../stats/saves/ex3_pragm_model_loglike.rds")
```

###### Noise parameter

```{r loglike rsa noise ex3, include = FALSE}
loglike.rsa.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

var model = function(){

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [1 - conditionInfo.prior, 1 - conditionInfo.prior, conditionInfo.prior]

      var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors, conditionInfo.speak_opt)

   var noisyModelPredictions = addNoise(modelPredictions, noise)

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )
  return  loglike

}, allParams)

 return sum(conditionOutput)

}
'
```

```{r}
# ex3_pragm_model_noise_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.rsa.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# ) %>%
#   saveRDS("../../stats/saves/ex3_pragm_model_noise_loglike.rds")

ex3_pragm_model_noise_loglike <- readRDS("../../stats/saves/ex3_pragm_model_noise_loglike.rds")%>%
  mutate(model = "pragmatic_noise")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))
```
##### Prior only model

###### Noise parameter

```{r loglike prior only noise ex3, include = FALSE}
loglike.prior.only.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

var model = function(){

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [1 - conditionInfo.prior, 1 - conditionInfo.prior, conditionInfo.prior]

  var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: conditionSpecificPriors}));
      return obj.shape == "circle" ? 1 : 0
  }})

   var noisyModelPredictions = addNoise(modelPredictions, noise)

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )
  return  loglike

}, allParams)

 return sum(conditionOutput)

}
'
```

```{r}
# ex3_prior_only_model_noise_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.prior.only.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# ) %>%
#   saveRDS("../../stats/saves/ex3_prior_only_model_noise_loglike.rds")

ex3_prior_only_model_noise_loglike <- readRDS("../../stats/saves/ex3_prior_only_model_noise_loglike.rds")%>%
  mutate(model = "prior_only_noise")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))
```

##### Flat prior model

###### Noise parameter

```{r loglike flat prior noise ex3, include = FALSE}
loglike.flat.prior.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

var model = function(){

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [.5, .5,.5]

      var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors, conditionInfo.speak_opt)

   var noisyModelPredictions = addNoise(modelPredictions, noise)

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )
  return  loglike

}, allParams)

 return sum(conditionOutput)

}
'
```

```{r}
# ex3_flat_prior_model_noise_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.flat.prior.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# ) %>%
#   saveRDS("../../stats/saves/ex3_flat_prior_model_noise_loglike.rds")

ex3_flat_prior_model_noise_loglike <- readRDS("../../stats/saves/ex3_flat_prior_model_noise_loglike.rds")%>%
  mutate(model = "flat_prior_noise")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))
```

```{r}
ex3_model_comparison <- bind_rows(
  log_like_pragmatic_ex3,
  ex3_pragm_model_noise_loglike,
  ex3_flat_prior_model_noise_loglike,
  ex3_prior_only_model_noise_loglike
)%>%
  mutate(model = reorder(model,logP))
```

#### Plot for likelihoods

```{r likelihood plot adults, include = T}
ggplot(ex3_model_comparison, aes(x = model, y = logP, col = model))+
  geom_segment( aes(x=model, xend=model, y=logP, yend=0), col = "black", size = 1)+
  geom_point(size = 7)+
  theme_few()+ 
  xlab("Model")+
  ylab("Log-likelihood")+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  guides(col = F)+
  scale_color_manual(breaks=c("pragmatic","pragmatic_noise","prior_only_noise","flat_prior_noise"),
                    values= c("#00afaf","#b58900","#859900","#6c71c4"))
```

#### Bayes factors

```{r bayes factors ex3 adults}

ex3_bf <- bind_rows(
  log_like_pragmatic_ex3,
  ex3_pragm_model_noise_loglike,
  ex3_flat_prior_model_noise_loglike,
  ex3_prior_only_model_noise_loglike
) %>% spread(model, logP) %>%
  mutate("pragmatic_noise > pragmatic" = exp(pragmatic_noise - pragmatic),
         "pragmatic_noise > prior_only_noise" = exp(pragmatic_noise - prior_only_noise),
         "pragmatic_noise > flat_prior_noise" = exp(pragmatic_noise - flat_prior_noise),
         "prior_only_noise > flat_prior_noise" = exp(prior_only_noise - flat_prior_noise)) %>%
  select(-pragmatic,-pragmatic_noise, -flat_prior_noise, -prior_only_noise)

ex3_bf %>%
  kable(digits = 2)
```

## Experiment 4

### Prior manipulation experiments

### Model parameters and data
```{r ex4 data adults}
# load data  from experiment 4 
adult_ex4_data <-read_csv(file="../../stats/data/adult_ex4.csv") 
```

```{r model setup ex 4}

# create datafiles with model parameters

ex4Prior <- expand.grid(
  common_ground_manipulation = c("preference","novelty"),
  speaker = c("same_speaker","different_speaker"),
  prior_manipulation = c("strong", "medium", "weak")
)

ex4_prior_summary <- adult_ex4_prior_data %>%
  group_by(common_ground_manipulation, prior_manipulation, speaker) %>%
  summarize(prior = mean(correct))

model_params_ex4 <- left_join( 
  ex4Prior, 
  ex4_prior_summary
) %>%
  rename(congruent = prior) %>%
  mutate(incongruent = 1 - congruent) %>%
  gather(alignment, prior, congruent, incongruent) %>%
  drop_na() %>%
  mutate(speak_opt = adult_model_param$so_map)
```

### Model predictions (including noise parameter)

#### Pragmatic model

```{r model rsa noise ex4, include = FALSE}
predictions.rsa.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

var model = function(){

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, prior_manipulation: conditionInfo.prior_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [1 - conditionInfo.prior, 1 - conditionInfo.prior, conditionInfo.prior]

      var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors, conditionInfo.speak_opt)

   var noisyModelPredictions = addNoise(modelPredictions, noise)

 map(function(d){ 
      observe(noisyModelPredictions, d.correct_inf)
    }, conditionSpecificData)

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.prior_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allParams)

  return extend(_.fromPairs(conditionOutput), {noise: noise})
}
'
```

```{r predictions pragmatic model ex 4 noise, include = FALSE}
# ex4_pragm_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex4.noise, sep='\n'),
#   data = list(allData = adult_ex4_data, allParams = model_params_ex4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_pragm_model_noise.rds")

# model predictions
ex4_pragm_model_pred_noise <- readRDS("../../stats/saves/ex4_pragm_model_noise.rds") %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation","prior_manipulation","speaker", "alignment"), sep="/")%>%
  mutate(model="pragmatic_noise")%>%
  group_by(model,common_ground_manipulation,prior_manipulation,speaker,alignment)%>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

## noise parameter
ex4_pragm_model_noise_param <- readRDS("../../stats/saves/ex4_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))
```

#### Prior only model

```{r model prior only noise ex4, include = FALSE}
predictions.prior.only.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

var model = function(){

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, prior_manipulation: conditionInfo.prior_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [1 - conditionInfo.prior, 1 - conditionInfo.prior, conditionInfo.prior]

  var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: conditionSpecificPriors}));
      return obj.shape == "circle" ? 1 : 0
  }})

   var noisyModelPredictions = addNoise(modelPredictions, noise)

 map(function(d){ 
      observe(noisyModelPredictions, d.correct_inf)
    }, conditionSpecificData)

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.prior_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allParams)

  return extend(_.fromPairs(conditionOutput), {noise: noise})
}
'
```

```{r predictions prior only model ex 4 noise, include = FALSE}
# ex4_prior_only_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.prior.only.ex4.noise, sep='\n'),
#   data = list(allData = adult_ex4_data, allParams = model_params_ex4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_prior_only_model_noise.rds")

# model predictions
ex4_prior_only_model_pred_noise <- readRDS("../../stats/saves/ex4_prior_only_model_noise.rds") %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation","prior_manipulation","speaker", "alignment"), sep="/")%>%
  mutate(model="prior_only_noise")%>%
  group_by(model,common_ground_manipulation,prior_manipulation,speaker,alignment)%>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

## noise parameter
ex4_prior_only_model_noise_param <- readRDS("../../stats/saves/ex4_prior_only_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))
```

#### Flat prior model

```{r model flat prior noise ex4, include = FALSE}
predictions.flat.prior.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

var model = function(){

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, prior_manipulation: conditionInfo.prior_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [.5, .5, .5]

      var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors, conditionInfo.speak_opt)

   var noisyModelPredictions = addNoise(modelPredictions, noise)

 map(function(d){ 
      observe(noisyModelPredictions, d.correct_inf)
    }, conditionSpecificData)

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.prior_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allParams)

  return extend(_.fromPairs(conditionOutput), {noise: noise})
}
'
```

```{r predictions flat prior model ex 4 noise, include = FALSE}
# ex4_flat_prior_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.flat.prior.ex4.noise, sep='\n'),
#   data = list(allData = adult_ex4_data, allParams = model_params_ex4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_flat_prior_model_noise.rds")

# model predictions
ex4_flat_prior_model_pred_noise <- readRDS("../../stats/saves/ex4_flat_prior_model_noise.rds") %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation","prior_manipulation","speaker", "alignment"), sep="/")%>%
  mutate(model="flat_prior_noise")%>%
  group_by(model,common_ground_manipulation,prior_manipulation,speaker,alignment)%>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

## noise parameter
ex4_flat_prior_model_noise_param <- readRDS("../../stats/saves/ex4_flat_prior_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))
```

```{r join model predictions and data ex4 adults}
# combine model predictions with data

# summarize data 
adult_ex4_data_summary <- adult_ex4_data %>%
  mutate(prior_manipulation = relevel(as.factor(prior_manipulation), ref = "strong"))%>%
  mutate(model = "data") %>%
  group_by(model,common_ground_manipulation,prior_manipulation,speaker,alignment)%>%
  summarize(k = sum(correct_inf), n = n())%>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         ci_lower  = qbeta(.025, a, b),
         ci_upper = qbeta(.975, a, b),
         mean = (a-1)/(a+b-2))%>%
  select(-a,-b,-n,-k)

# summarize model predictions
ex_4_model_pred <- bind_rows(
  ex4_pragm_model_pred_noise,
  ex4_prior_only_model_pred_noise,
  ex4_flat_prior_model_pred_noise
)%>%
  ungroup()%>%
  mutate(prior_manipulation = relevel(as.factor(prior_manipulation), ref = "strong"))

# combine model predictions and data

ex_4_pred <- bind_rows(
  ex_4_model_pred,
  adult_ex4_data_summary
)

#saveRDS(ex_4_pred, "../../stats/saves/adult_ex4_model_predictions_data.rds")

```

### Compare model predictions to data

##### Noise parameters

```{r, include = T}
ex4_noise_parameters <- bind_rows(
  readRDS("../../stats/saves/ex4_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  mutate(model = "pragmatic"),
  readRDS("../../stats/saves/ex4_prior_only_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  mutate(model = "prior_only"),
  readRDS("../../stats/saves/ex4_flat_prior_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  mutate(model = "flat_prior")
  )

ggplot(ex4_noise_parameters, aes(x = value, col = model, fill = model))+
  geom_density(alpha = .5)+
  theme_few()+
  xlim(0,1)+
  scale_color_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#b58900","#6c71c4","#859900"))+
  scale_fill_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#b58900","#6c71c4","#859900"))
```

#### Plot model predictions and data

```{r plot ex4 model predictions and data adults, include = T}
ggplot(data = ex_4_pred, aes(x = alignment, y = mean, col = model)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), size = .4,  position = position_dodge(width = .5), pch = 5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion more informative")+
  facet_grid ( prior_manipulation ~ common_ground_manipulation + speaker) +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_color_manual(name="Model",
                     breaks=c("data","pragmatic_noise","flat_prior_noise","prior_only_noise"),
                     values= c("black","#b58900","#6c71c4","#859900"))+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))

```

#### Correlation plot model predictions and data

```{r correlation plot ex4 adults, include = T}
ex_4_cor_plot <- ex_4_pred %>%
  ungroup()%>%
  filter(model != "data") %>%
  left_join(., ex_4_pred %>%
    ungroup()%>%
    filter(model == "data") %>%
    rename(data_mean = mean, data_ci_lower = ci_lower, data_ci_upper = ci_upper) %>%
    select(-model)
  )

ggplot(data = ex_4_cor_plot,aes(x = mean, y = data_mean, col = prior_manipulation)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 0.7, size = 0.5)+
  geom_point(size = 2)+
  geom_errorbar(aes(ymin = data_ci_lower, ymax = data_ci_upper),width = 0,size = .7)+
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0,size = .7)+
  coord_fixed()+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  ylab("Data")+
  facet_grid ( ~ model) +
  stat_cor(method = "pearson", label.x = 0.01, label.y = 0.99, aes(x = mean, y = data_mean), inherit.aes = F, size = 3)+
  theme_few(base_size = ) + 
  scale_colour_ptol()
```

### Model comparison

#### Marginal likelihoods

##### Pragmatic model 
```{r loglike rsa noise ex4, include = FALSE}
loglike.rsa.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

var model = function(){

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

 var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, prior_manipulation: conditionInfo.prior_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [1 - conditionInfo.prior, 1 - conditionInfo.prior, conditionInfo.prior]

      var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors, conditionInfo.speak_opt)

   var noisyModelPredictions = addNoise(modelPredictions, noise)

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )
  return  loglike

}, allParams)

 return sum(conditionOutput)

}
'
```

```{r}
# ex4_pragm_model_noise_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.rsa.ex4.noise, sep='\n'),
#   data = list(allData = adult_ex4_data, allParams = model_params_ex4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# ) %>%
#   saveRDS("../../stats/saves/ex4_pragm_model_noise_loglike.rds")

ex4_pragm_model_noise_loglike <- readRDS("../../stats/saves/ex4_pragm_model_noise_loglike.rds")%>%
  mutate(model = "pragmatic_noise")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))
```

##### Prior only model

```{r loglike prior only noise ex4, include = FALSE}
loglike.prior.only.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

var model = function(){

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

 var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, prior_manipulation: conditionInfo.prior_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [1 - conditionInfo.prior, 1 - conditionInfo.prior, conditionInfo.prior]

  var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: conditionSpecificPriors}));
      return obj.shape == "circle" ? 1 : 0
  }})

   var noisyModelPredictions = addNoise(modelPredictions, noise)

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )
  return  loglike

}, allParams)

 return sum(conditionOutput)

}
'
```

```{r}
# ex4_prior_only_model_noise_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.prior.only.ex4.noise, sep='\n'),
#   data = list(allData = adult_ex4_data, allParams = model_params_ex4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# ) %>%
#   saveRDS("../../stats/saves/ex4_prior_only_model_noise_loglike.rds")

ex4_prior_only_model_noise_loglike <- readRDS("../../stats/saves/ex4_prior_only_model_noise_loglike.rds")%>%
  mutate(model = "prior_only_noise")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))
```

##### Flat prior model

```{r loglike flat prior noise ex4, include = FALSE}
loglike.flat.prior.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

var model = function(){

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

 var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, prior_manipulation: conditionInfo.prior_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [.5, .5,.5]

      var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors, conditionInfo.speak_opt)

   var noisyModelPredictions = addNoise(modelPredictions, noise)

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )
  return  loglike

}, allParams)

 return sum(conditionOutput)

}
'
```

```{r}
# ex4_flat_prior_model_noise_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.flat.prior.ex4.noise, sep='\n'),
#   data = list(allData = adult_ex4_data, allParams = model_params_ex4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# ) %>%
#   saveRDS("../../stats/saves/ex4_flat_prior_model_noise_loglike.rds")

ex4_flat_prior_model_noise_loglike <- readRDS("../../stats/saves/ex4_flat_prior_model_noise_loglike.rds")%>%
  mutate(model = "flat_prior_noise")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))
```

```{r}
ex4_model_comparison <- bind_rows(
  ex4_pragm_model_noise_loglike,
  ex4_flat_prior_model_noise_loglike,
  ex4_prior_only_model_noise_loglike
)%>%
  mutate(model = reorder(model,logP))
```

#### Plot for likelihoods

```{r plot for loglikelihood ex4 adults, include = T}
ggplot(ex4_model_comparison, aes(x = model, y = logP, col = model))+
  geom_segment( aes(x=model, xend=model, y=logP, yend=0), col = "black", size = 1)+
  geom_point(size = 7)+
  theme_few()+ 
  xlab("Model")+
  ylab("Log-likelihood")+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  guides(col = F)+
  scale_color_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#859900","#b58900","#6c71c4"))
```

#### Bayes factors

```{r bayes factors ex4 adults, include = T}
ex4_bf <- bind_rows(
  ex4_pragm_model_noise_loglike,
  ex4_flat_prior_model_noise_loglike,
  ex4_prior_only_model_noise_loglike
) %>% spread(model, logP) %>%
  mutate("pragmatic_noise > flat_prior_noise" = exp(pragmatic_noise - flat_prior_noise),
         "pragmatic_noise > prior_only_noise" = exp(pragmatic_noise - prior_only_noise),
         "flat_prior_noise > prior_only_noise" = exp(flat_prior_noise - prior_only_noise)) %>%
  select(-pragmatic_noise,  -flat_prior_noise, -prior_only_noise)

ex4_bf %>%
  kable(digits = 2)
```

# Children

## Model parameters and data

```{r ex3 data children}
# children ex3 data
child_ex3_data <-read_csv(file="../../stats/data/child_ex3.csv") 
```

```{r ex1 and ex2 data children}
# load data from ex1 manipulation experiments to set speaker optimality parameter
child_ex1_data <-read_csv(file="../../stats/data/child_ex1.csv")
# load data from ex2 experiment to set priors
child_ex2_data <-read_csv(file="../../stats/data/child_ex2.csv") 
```

#### Inferring speaker optimality parameter and priors

```{r model parameters children}
# The prior on the speaker optimality intercept is the same as for the adult speaker optimality parameter.

childModelParam <- '
var prefData = dataFromR.prefData
var infData = dataFromR.infData

var all_conditions = levels(prefData, "condition")

var model  = function(){

  var so_slope = sample(Cauchy({location: 0, scale: 2}))
  var so_int = sample(Cauchy({location: 1, scale: 2}))

  foreach(function(row){

  var age = row.age_num
  var speakerOptimality = so_int  + so_slope * (age - infData[0].minage)

    var inf_priorProbs = [.5, .5, .5]

  var rsaPredictions = pragmaticListener({label: "dax", point: 2 },
   inf_priorProbs, speakerOptimality) 
   observe(rsaPredictions, row.correct)

  }, infData)

  var pref_params = map(function(cndtn){

    var conditionData = _.filter(prefData, {condition: cndtn})

    var pref_slope = uniformDrift({a: -2, b: 2, width: 0.4})
    var pref_int = uniformDrift({a: -2, b: 2, width: 0.4})

    foreach(function(row){

      var age = row.age_num

      var priorReg = logistic(pref_int + pref_slope * (age - row.minage))
      var priorProbs= [1-priorReg, priorReg]

      var priorPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: prior_objects, ps: priorProbs}));
      return obj.shape == "circle" ? 1 : 0
      }})

      observe(priorPredictions, row.correct)

    }, conditionData)

    return [
      {param: "pref_int", val: pref_int, condition: cndtn}, 
      {param: "pref_slope", val: pref_slope, condition: cndtn}
    ]

  }, all_conditions)

  return _.flatten(pref_params).concat(
    [
      { param: "so_slope", val: so_slope},
      { param: "so_int", val: so_int}
    ]
  )
}
'
```

```{r inferring model parameters children}
# child_model_param<- webppl(
#   program_code = paste(rsaUtils, rsaModel, childModelParam , sep='\n'),
#   data = list(prefData = child_ex2_data, infData = child_ex1_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# )%>%
#   saveRDS("../../stats/saves/child_model_param.rds")

child_model_param <- readRDS("../../stats/saves/child_model_param.rds")%>%
  select(value) %>%
  map_df(bind_rows)%>%
  group_by(param,condition)%>%
  summarise(param_map = estimate_mode(val))%>%
  mutate(cond = paste(param,condition, sep = "_"))%>%
  ungroup()%>%
  select(cond,param_map)%>%
  spread(cond,param_map)%>%
  rename(so_int = so_int_NA,
         so_slope = so_slope_NA)

```

```{r visualize model parameters children, include = T}
# visualize inferred parameter
plot_child_model_param <- readRDS("../../stats/saves/child_model_param.rds")%>%
  select(value)%>%
  map_df(bind_rows)%>%
  mutate(type = ifelse(grepl("pref",param),"prior", "speaker optimality"))

ggplot(plot_child_model_param, aes(val))+
  geom_density()+
  xlab("Parameter")+
  facet_grid(type~param+condition)+
  theme_few()
```

```{r visualize prior children, include = T}
# visualize model priors in comparison to data 
child_prior_plot <- child_ex2_data %>%
  group_by(id,age_num,minage,condition)%>%
  summarise(correct = mean(correct))%>%
  mutate(model = ifelse(condition == "different_speaker",
                        plogis(child_model_param$pref_int_different_speaker + child_model_param$pref_slope_different_speaker * (age_num-minage)),
                        plogis(child_model_param$pref_int_same_speaker + child_model_param$pref_slope_same_speaker * (age_num-minage)
                        )))

ggplot(data = child_prior_plot, aes(x = age_num, y = correct, col = condition)) +
  geom_jitter(alpha = 0.5, height = 0.02, width = 0)+
  geom_smooth(aes(fill = condition),method = "lm", se = T, alpha = .5, size = .7)+
  geom_line(aes(x = age_num, y = model, col = condition), size = 1.5, lty = 2)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="Age",y="Proportion Correct")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_color_solarized(name = "Condition")+
  scale_fill_solarized(name = "Condition")

# dashed line represents the developmental trajectories based on the estimated parameters 
```

```{r visualize speaker optimality children, include = T}
# visualize speaker optimality in comparison to data 
child_so_plot <- child_ex1_data %>%
  group_by(id,age_num,minage)%>%
  summarise(correct = mean(correct))%>%
  mutate(model = child_model_param$so_int + child_model_param$so_slope * (age_num - minage))

plot_child_so <- ggplot(data = child_so_plot, aes(x = age_num, y = model)) +
  geom_hline(yintercept = 0, lty=2)+
  geom_line( size = 2)+
  labs(x="Age",y="Speaker optimality parameter")+
  theme_few() +
  ggtitle("Speaker optimality parameter")+
  ylim(-3,3)+
  guides(alpha = F)

plot_ex1_child <-child_ex1_data %>%
  group_by(age_num, id) %>%
  summarise(correct = mean(correct))%>%
  ggplot(data = ., aes(x = age_num, y = correct)) +
  geom_jitter(alpha = 0.5, height = 0.02, width = 0)+
  geom_smooth(method = "lm", se = T, alpha = .5, size = .7)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="Age",y="Proportion Correct")+
  ggtitle("Data ex1 children")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)

ggarrange(plot_child_so,plot_ex1_child)
```

```{r model setup ex3 children}
ex3_model_data <- child_ex3_data%>%
  mutate(so_int = child_model_param$so_int,
         so_slope = child_model_param$so_slope,
         pref_int_different_speaker = child_model_param$pref_int_different_speaker,
         pref_slope_different_speaker = child_model_param$pref_slope_different_speaker,
         pref_int_same_speaker = child_model_param$pref_int_same_speaker,
         pref_slope_same_speaker = child_model_param$pref_slope_same_speaker,
         so = so_int+so_slope*(age_num-min(age_num)),
         diff_prior = plogis(pref_int_different_speaker+pref_slope_different_speaker*(age_num-min(age_num))),
         same_prior = plogis(pref_int_same_speaker+pref_slope_same_speaker*(age_num-min(age_num))))%>%
  group_by(id,trial)%>%
        mutate(priorProbs = ifelse(speaker == "same_speaker" & alignment == "congruent", 
                                   list(c(1-same_prior,1-same_prior,same_prior)),
                                   ifelse(speaker == "different_speaker" & alignment == "congruent",
                                          list(c(1-diff_prior,1-diff_prior,diff_prior)),
                                          ifelse(speaker == "same_speaker" & alignment == "incongruent",
                                                 list(c(same_prior,same_prior,1-same_prior)),
                                                 list(c(diff_prior,diff_prior,1-diff_prior))
                                   ))))%>%
  ungroup()%>%
  select(speaker,age_num,minage,alignment,so,priorProbs,correct_inf)

```
## Model predictions

### Pragmatic model

#### Parameter free

```{r pragmatic model ex3 children}
child.ex3.rsa.predictions <-'

var allData = dataFromR.allData

var output = map(function(row){

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, row.priorProbs,row.so)

    return extend([row.age_num + "/" + row.speaker + "/" + row.alignment + "/" + row.correct_inf, Math.exp(modelPredictions.score(1))])

}, allData)

output

'
```

```{r predictions pragmatic model ex3 children}
child_ex3_pragm_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, child.ex3.rsa.predictions , sep='\n'),
  data =list(allData = ex3_model_data),
  data_var = "dataFromR"
)

# model predictions
child_ex3_pragm_model_predictions <- child_ex3_pragm_model %>%
  separate(`0`, into = c("age", "speaker", "alignment","correct"), sep="/")%>%
  mutate(mean = `1`,
         ci_lower = mean,
         ci_upper = mean,
         model="pragmatic",
         age = as.numeric(age))%>%
  select(-`1`)
```

#### Noise parameter

```{r}
child.ex3.rsa.predictions.noise <-'

var allData = dataFromR.allData

var model = function(){

  var noise = uniformDrift({a: 0, b: 1, width: 0.1})

  var output = map(function(row){

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, row.priorProbs,row.so)

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct_inf)

    return  [row.age_num + "/" + row.speaker + "/" + row.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allData)

 return extend(_.fromPairs(output), {noise: noise})

}
'
```

```{r}
# child_ex3_pragm_model_noise<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.rsa.predictions.noise , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# )%>%
#    saveRDS("../../stats/saves/child_ex3_pragm_model_noise.rds")

# model predictions
child_ex3_pragm_model_predictions_noise <- readRDS("../../stats/saves/child_ex3_pragm_model_noise.rds") %>%
  filter(!(Parameter %in% c("noise"))) %>%
  separate(Parameter, into = c("age", "speaker", "alignment"), sep="/")%>%
  mutate(model = "pragmatic_noise")%>%
  group_by(model, age,alignment,speaker) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))

## noise parameter
child_ex3_pragm_model_noise_param <- readRDS("../../stats/saves/child_ex3_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

```

#### Developmental noise parameter

```{r}
child.ex3.rsa.predictions.developmental.noise <-'

var allData = dataFromR.allData

var model = function(){

  var noise_slope = uniformDrift({a: -5, b: 5, width: 0.2})
  var noise_int = uniformDrift({a: -5, b: 5, width: 0.1})

  var output = map(function(row){

    var noise = logistic(noise_int + noise_slope * (row.age_num-row.minage))

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, row.priorProbs,row.so)

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct_inf)

    return  [row.age_num + "/" + row.speaker + "/" + row.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allData)

 return extend(_.fromPairs(output), {noise_int: noise_int, noise_slope: noise_slope})

}
'
```

```{r}
# child_ex3_pragm_model_developmental_noise<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.rsa.predictions.developmental.noise , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# )%>%
#    saveRDS("../../stats/saves/child_ex3_pragm_model_developmental_noise.rds")

# model predictions
child_ex3_pragm_model_predictions_developmental_noise <- readRDS("../../stats/saves/child_ex3_pragm_model_developmental_noise.rds") %>%
  filter(!(Parameter %in% c("noise","noise_int","noise_slope"))) %>%
  separate(Parameter, into = c("age", "speaker", "alignment"), sep="/")%>%
  mutate(model = "pragmatic_developmental_noise")%>%
  group_by(model, age,alignment,speaker) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))

## noise parameter
child_ex3_pragm_model_developmental_noise_param <- readRDS("../../stats/saves/child_ex3_pragm_model_developmental_noise.rds") %>%
  filter(Parameter %in% c("noise_int","noise_slope"))  %>%
  group_by(Parameter)%>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

```

### Prior only model

#### Parameter free

```{r prior only model ex3 children}
child.ex3.prior.only.predictions <-'

var allData = dataFromR.allData

  var output = map(function(row){

      var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: row.priorProbs}));
      return obj.shape == "circle" ? 1 : 0
       }})

    return  extend([row.age_num + "/" + row.speaker + "/" + row.alignment+ "/" + row.correct_inf, Math.exp(modelPredictions.score(1))])

}, allData)

 output

'
```

```{r predictions prior only model ex3 children}
child_ex3_prior_only_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, child.ex3.prior.only.predictions , sep='\n'),
  data =list(allData = ex3_model_data),
  data_var = "dataFromR"
)

# model predictions
child_ex3_prior_only_model_predictions <- child_ex3_prior_only_model %>%
  separate(`0`, into = c("age", "speaker", "alignment","correct"), sep="/")%>%
  mutate(mean = `1`,
         model="prior_only",
         age = as.numeric(age))%>%
  select(-`1`)

```

#### Noise parameter

```{r}
child.ex3.prior.only.predictions.noise <-'

var allData = dataFromR.allData

var model = function(){

  var noise = uniformDrift({a: 0, b: 1, width: 0.1})

  var output = map(function(row){

      var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: row.priorProbs}));
      return obj.shape == "circle" ? 1 : 0
       }})

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct_inf)

    return  [row.age_num + "/" + row.speaker + "/" + row.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allData)

 return extend(_.fromPairs(output), {noise: noise})

}
'
```

```{r}
# child_ex3_prior_only_model_noise<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.prior.only.predictions.noise , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# )%>%
#    saveRDS("../../stats/saves/child_ex3_prior_only_model_noise.rds")

# model predictions
child_ex3_prior_only_model_predictions_noise <- readRDS("../../stats/saves/child_ex3_prior_only_model_noise.rds") %>%
  filter(!(Parameter %in% c("noise"))) %>%
  separate(Parameter, into = c("age", "speaker", "alignment"), sep="/")%>%
  mutate(model = "prior_only_noise")%>%
  group_by(model, age,alignment,speaker) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))

## noise parameter
child_ex3_prior_only_model_noise_param <- readRDS("../../stats/saves/child_ex3_prior_only_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

```

#### Developmental noise parameter

```{r}
child.ex3.prior.only.predictions.developmental.noise <-'

var allData = dataFromR.allData

var model = function(){

  var noise_slope = uniformDrift({a: -5, b: 5, width: 0.2})
  var noise_int = uniformDrift({a: -5, b: 5, width: 0.1})

  var output = map(function(row){

    var noise = logistic(noise_int + noise_slope * (row.age_num-row.minage))

      var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: row.priorProbs}));
      return obj.shape == "circle" ? 1 : 0
       }})

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct_inf)

    return  [row.age_num + "/" + row.speaker + "/" + row.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allData)

 return extend(_.fromPairs(output), {noise_int: noise_int, noise_slope: noise_slope})

}
'
```

```{r}
# child_ex3_prior_only_model_developmental_noise<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.prior.only.predictions.developmental.noise , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# )%>%
#    saveRDS("../../stats/saves/child_ex3_prior_only_model_developmental_noise.rds")

# model predictions
child_ex3_prior_only_model_predictions_developmental_noise <- readRDS("../../stats/saves/child_ex3_prior_only_model_developmental_noise.rds") %>%
  filter(!(Parameter %in% c("noise_int","noise_slope"))) %>%
  separate(Parameter, into = c("age", "speaker", "alignment"), sep="/")%>%
  mutate(model = "prior_only_developmental_noise")%>%
  group_by(model, age,alignment,speaker) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))

## noise parameter
child_ex3_prior_only_model_developmental_noise_param <- readRDS("../../stats/saves/child_ex3_prior_only_model_developmental_noise.rds") %>%
  filter(Parameter %in% c("noise_int","noise_slope"))  %>%
  group_by(Parameter)%>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

```

### Flat prior model

#### Parameter free

```{r flat prior model ex3 children}
child.ex3.flat.prior.predictions <-'

var allData = dataFromR.allData

  var output = map(function(row){

    var prior = [.5, .5, .5]

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, prior,row.so)

    return  extend([row.age_num + "/" + row.speaker + "/" + row.alignment + "/" + row.correct_inf, Math.exp(modelPredictions.score(1))])

}, allData)

output

'
```

```{r predictions flat prior model ex3 children}
child_ex3_flat_prior_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, child.ex3.flat.prior.predictions , sep='\n'),
  data =list(allData = ex3_model_data),
  data_var = "dataFromR")

# model predictions
child_ex3_flat_prior_model_predictions <- child_ex3_flat_prior_model %>%
  separate(`0`, into = c("age", "speaker", "alignment","correct"), sep="/")%>%
  mutate(mean = `1`,
         model="flat_prior",
         age = as.numeric(age))%>%
  select(-`1`)

```

#### Noise parameter

```{r}
child.ex3.flat.prior.predictions.noise <-'

var allData = dataFromR.allData

var model = function(){

  var noise = uniformDrift({a: 0, b: 1, width: 0.1})

  var output = map(function(row){

    var prior = [.5,.5,.5]

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, prior,row.so)

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct_inf)

    return  [row.age_num + "/" + row.speaker + "/" + row.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allData)

 return extend(_.fromPairs(output), {noise: noise})

}
'
```

```{r}
# child_ex3_flat_prior_model_noise<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.flat.prior.predictions.noise , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# )%>%
#    saveRDS("../../stats/saves/child_ex3_flat_prior_model_noise.rds")

# model predictions
child_ex3_flat_prior_model_predictions_noise <- readRDS("../../stats/saves/child_ex3_flat_prior_model_noise.rds") %>%
  filter(!(Parameter %in% c("noise"))) %>%
  separate(Parameter, into = c("age", "speaker", "alignment"), sep="/")%>%
  mutate(model = "flat_prior_noise")%>%
  group_by(model, age,alignment,speaker) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))

## noise parameter
child_ex3_flat_prior_model_noise_param <- readRDS("../../stats/saves/child_ex3_flat_prior_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

```

#### Developmental noise parameter

```{r}
child.ex3.flat.prior.predictions.developmental.noise <-'

var allData = dataFromR.allData

var model = function(){

  var noise_slope = uniformDrift({a: -5, b: 5, width: 0.2})
  var noise_int = uniformDrift({a: -5, b: 5, width: 0.1})

  var output = map(function(row){

    var noise = logistic(noise_int + noise_slope * (row.age_num-row.minage))

    var prior = [.5,.5,.5]

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, prior,row.so)

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct_inf)

    return  [row.age_num + "/" + row.speaker + "/" + row.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allData)

 return extend(_.fromPairs(output), {noise_int: noise_int, noise_slope: noise_slope})

}
'
```

```{r}
# child_ex3_flat_prior_model_developmental_noise<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.flat.prior.predictions.developmental.noise , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# )%>%
#    saveRDS("../../stats/saves/child_ex3_flat_prior_model_developmental_noise.rds")

# model predictions
child_ex3_flat_prior_model_predictions_developmental_noise <- readRDS("../../stats/saves/child_ex3_flat_prior_model_developmental_noise.rds") %>%
  filter(!(Parameter %in% c("noise","noise_int","noise_slope"))) %>%
  separate(Parameter, into = c("age", "speaker", "alignment"), sep="/")%>%
  mutate(model = "flat_prior_developmental_noise")%>%
  group_by(model, age,alignment,speaker) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))

## noise parameter
child_ex3_flat_prior_model_developmental_noise_param <- readRDS("../../stats/saves/child_ex3_flat_prior_model_developmental_noise.rds") %>%
  filter(Parameter %in% c("noise_int","noise_slope"))  %>%
  group_by(Parameter)%>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

```

## Compare model predictions to data

### Noise parameters

```{r, include = T}
child_noise_parameters <- bind_rows(
  readRDS("../../stats/saves/child_ex3_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  mutate(model = "pragmatic"),
  readRDS("../../stats/saves/child_ex3_prior_only_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  mutate(model = "prior_only"),
  readRDS("../../stats/saves/child_ex3_flat_prior_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  mutate(model = "flat_prior")
  )

ggplot(child_noise_parameters, aes(x = value, col = model, fill = model))+
  geom_density(alpha = .5)+
  theme_few()+
  xlim(0,1)+
  scale_color_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#b58900","#6c71c4","#859900"))+
  scale_fill_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#b58900","#6c71c4","#859900"))
```

### Developmental noise parameters
```{r, include = T}
developmental_noise_parameter <- data_frame(
  age = rep(seq(0,2, by = 0.05),3),
  model = c(rep("pragmatic",41),
            rep("flat_prior",41),
            rep("prior_only",41)),
  int = c(rep(child_ex3_pragm_model_developmental_noise_param%>%
                filter(Parameter == "noise_int")%>%
                pull(mean),41),
          rep(child_ex3_flat_prior_model_developmental_noise_param%>%
                filter(Parameter == "noise_int")%>%
                pull(mean),41),
          rep(child_ex3_prior_only_model_developmental_noise_param%>%
                filter(Parameter == "noise_int")%>%
                pull(mean),41)
          ),
  slope = c(rep(child_ex3_pragm_model_developmental_noise_param%>%
                filter(Parameter == "noise_slope")%>%
                pull(mean),41),
          rep(child_ex3_flat_prior_model_developmental_noise_param%>%
                filter(Parameter == "noise_slope")%>%
                pull(mean),41),
          rep(child_ex3_prior_only_model_developmental_noise_param%>%
                filter(Parameter == "noise_slope")%>%
                pull(mean),41))
) %>%
  mutate(y = plogis(int + slope *age))

ggplot(developmental_noise_parameter, aes(x = age, y= y, col = model))+
  geom_line(size = 1)+
  ylab("Noise parameter")+
  xlab("Age")+
  ylim(0,1)+
  theme_few()+
  scale_color_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#b58900","#6c71c4","#859900"))
```

### Plot model predictions and data

```{r plot model predictions ex3 children, include = T}
child_model_predictions <- bind_rows(
  child_ex3_pragm_model_predictions_noise,
  child_ex3_pragm_model_predictions_developmental_noise, 
  child_ex3_pragm_model_predictions
  #child_ex3_prior_only_model_predictions_noise,
  #child_ex3_flat_prior_model_predictions_noise
  #child_ex3_flat_prior_model_predictions,
  #child_ex3_prior_only_model_predictions
  )

#saveRDS(child_model_predictions, "../../stats/saves/child_ex3_model_predictions.rds")

plot_child_ex3_data <- child_ex3_data%>%
  group_by(age_num,speaker,alignment)%>%
  summarise(mean = mean(correct_inf))

ggplot(child_model_predictions)+
  geom_jitter(data = plot_child_ex3_data, aes(x = age_num, y = mean), width = .025,height = .025, alpha = .2)+
  geom_ribbon(aes(x = age, ymin = ci_lower, ymax = ci_upper, fill = model), alpha = .4) +
  geom_line(aes(x = age, y = mean, col = model), size = 1, alpha = 1, linetype = 1) + 
  #geom_smooth(data = child_ex3_data, aes(x = age_num, y = correct_inf), col = "black", size = 0.6, method = "loess", span = 2,  se = T,  alpha = .5)+
  #geom_hline(yintercept = 0.5, lty=2)+
  labs(y="Proportion more informative")+
  ylim(-0.05,1.05)+
  facet_grid(~speaker+alignment)+
  theme_few()+
  theme(legend.position="bottom")+
  scale_color_manual(name="Model",
                     breaks=c("pragmatic","pragmatic_noise","pragmatic_developmental_noise"),
                     values= c("#00afaf","#af6300","#6c71c4"))+
  scale_fill_manual(name="Model",
                     breaks=c("pragmatic","pragmatic_noise","pragmatic_developmental_noise"),
                     values= c("#00afaf","#af6300","#6c71c4"))

```

### Correlation plot model predictions and data

```{r correlation plot ex3 children, include = T}
child_cor_data <- child_ex3_data %>%
  group_by(age_bin,speaker, alignment) %>%
  summarize(k = sum(correct_inf), n = n())%>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         data_ci_lower  = qbeta(.025, a, b),
         data_ci_upper = qbeta(.975, a, b),
         data_mode = (a-1)/(a+b-2))%>%
  select(-a,-b,-n,-k)

child_cor_model <- bind_rows(
  child_ex3_pragm_model %>%
  separate(`0`, into = c("age", "speaker", "alignment","correct"), sep="/")%>%
  mutate(value = `1`,
         model = "pragmatic_parameter_free",
         age = ifelse(age<4,"3","4"))%>%
    group_by(age,model,speaker,alignment)%>%
    summarise(model_mode = estimate_mode(value),
              model_ci_lower = hdi_lower(value), 
              model_ci_upper =hdi_upper(value)),
  readRDS("../../stats/saves/child_ex3_pragm_model_noise.rds") %>%
    filter(!(Parameter %in% c("noise"))) %>%
    separate(Parameter, into = c("age", "speaker", "alignment"), sep="/")%>%
    mutate(model = "pragmatic_noise",
           age = ifelse(age<4,"3","4"))%>%
    group_by(age,model,speaker,alignment) %>%
    summarise(model_mode = estimate_mode(value), 
              model_ci_lower = hdi_lower(value), 
              model_ci_upper =hdi_upper(value)),
  readRDS("../../stats/saves/child_ex3_pragm_model_developmental_noise.rds") %>%
    filter(!(Parameter %in% c("noise","noise_int","noise_slope"))) %>%
    separate(Parameter, into = c("age", "speaker", "alignment"), sep="/")%>%
    mutate(model = "pragmatic_developmental_noise",
           age = ifelse(age<4,"3","4"))%>%
    group_by(age,model,speaker,alignment) %>%
    summarise(model_mode = estimate_mode(value), 
              model_ci_lower = hdi_lower(value), 
              model_ci_upper =hdi_upper(value))
)

child_cor_plot <- child_cor_model %>%
  ungroup()%>%
  mutate(data_mode = rep(child_cor_data$data_mode,3),
         data_ci_lower = rep(child_cor_data$data_ci_lower,3),
         data_ci_upper  =rep( child_cor_data$data_ci_upper,3))

ggplot(data = child_cor_plot,aes(x = model_mode, y = data_mode, col = age)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_point(size = 3)+
  geom_errorbar(aes(ymin = data_ci_lower, ymax = data_ci_upper),width = 0,size = .7)+
  geom_errorbarh(aes(xmin = model_ci_lower, xmax = model_ci_upper), height = 0,size = .7)+
  coord_fixed()+
  facet_grid(~model)+
  stat_cor(method = "pearson", label.x = 0.01, label.y = 0.99, aes(x = model_mode, y = data_mode), inherit.aes = F)+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  ylab("Data")+
  theme_few() + 
  scale_colour_ptol()
```

## Model comparison

### Marginal likelihoods

#### Pragmatic model

##### Parameter free

```{r loglike pragmatic model ex3 children}
child.ex3.rsa.marginal.likelihood <-'

var allData = dataFromR.allData

var model = function(){

  var output = map(function(row){

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, row.priorProbs,row.so)

    observe(modelPredictions, row.correct_inf)

}, allData)

}

var steps = 5000
var samples = 1
var mll = AIS(model, {steps:steps, samples: samples})
mll

'
```

```{r}
# child_ex3_pragm_model_likelihood<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.rsa.marginal.likelihood , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   chains = 3,
#   cores = 3
# )%>%
#   saveRDS("../../stats/saves/child_ex3_pragm_model_loglike.rds")

child_ex3_pragm_model_likelihood <- readRDS("../../stats/saves/child_ex3_pragm_model_loglike.rds")%>%
  as.tibble()%>%
  mutate(model = "pragmatic",
         parameter = "parameter free")%>%
  group_by(model,parameter)%>%
  summarise(logP = mean(V1))

```

##### Noise parameter

```{r loglike pragmatic model noise ex3 children}
child.ex3.rsa.noise.marginal.likelihood <-'

var allData = dataFromR.allData

var model = function(){

  var noise = uniform({a: 0, b: 1})

  var output = map(function(row){

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, row.priorProbs,row.so)

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct_inf)

}, allData)

}

var steps = 5000
var samples = 1
var mll = AIS(model, {steps:steps, samples: samples})
mll
'
```

```{r}
# child_ex3_pragm_model_noise_likelihood<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.rsa.noise.marginal.likelihood , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   chains = 3,
#   cores = 3
# )%>%
#   saveRDS("../../stats/saves/child_ex3_pragm_model_noise_loglike.rds")

child_ex3_pragm_model_noise_likelihood <- readRDS("../../stats/saves/child_ex3_pragm_model_noise_loglike.rds")%>%
  as.tibble()%>%
  mutate(model = "pragmatic",
         parameter = "noise")%>%
  group_by(model,parameter)%>%
  summarise(logP = mean(V1))

```

##### Developmental noise parameter

```{r loglike pragmatic model developmental noise ex3 children}
child.ex3.rsa.developmental.noise.marginal.likelihood <-'

var allData = dataFromR.allData

var model = function(){

  var noise_slope = uniformDrift({a: -5, b: 5, width: 0.2})
  var noise_int = uniformDrift({a: -5, b: 5, width: 0.1})

  var output = map(function(row){

    var noise = logistic(noise_int + noise_slope * (row.age_num-row.minage))

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, row.priorProbs,row.so)

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct_inf)

}, allData)

}

var steps = 5000
var samples = 1
var mll = AIS(model, {steps:steps, samples: samples})
mll
'
```

```{r}
# child_ex3_pragm_model_developmental_noise_likelihood<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.rsa.developmental.noise.marginal.likelihood , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   chains = 3,
#   cores = 3
# )%>%
#   saveRDS("../../stats/saves/child_ex3_pragm_model_developmenatl_noise_loglike.rds")

child_ex3_pragm_model_developmental_noise_likelihood <- readRDS("../../stats/saves/child_ex3_pragm_model_developmenatl_noise_loglike.rds")%>%
  as.tibble()%>%
  mutate(model = "pragmatic",
         parameter = "developmental noise")%>%
  group_by(model,parameter)%>%
  summarise(logP = mean(V1))

```

#### Flat prior model

##### Parameter free

```{r}
child.ex3.flat.prior.marginal.likelihood <-'

var allData = dataFromR.allData

var model = function(){

  var noise = uniform({a: 0, b: 1})

  var output = map(function(row){

    var prior = [.5,.5,.5]

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, prior,row.so)

    observe(modelPredictions, row.correct_inf)

}, allData)

}

var steps = 5000
var samples = 1
var mll = AIS(model, {steps:steps, samples: samples})
mll
'
```

```{r}
# child_ex3_flat_prior_model_likelihood<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.flat.prior.marginal.likelihood , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   chains = 3,
#   cores = 3
# )%>%
#   saveRDS("../../stats/saves/child_ex3_flat_prior_model_loglike.rds")

child_ex3_flat_prior_model_likelihood <- readRDS("../../stats/saves/child_ex3_flat_prior_model_loglike.rds")%>%
  as.tibble()%>%
  mutate(model = "flat_prior",
         parameter = "parameter free")%>%
  group_by(model,parameter)%>%
  summarise(logP = mean(V1))
```

##### Noise parameter

```{r}
child.ex3.flat.prior.noise.marginal.likelihood <-'

var allData = dataFromR.allData

var model = function(){

  var noise = uniform({a: 0, b: 1})

  var output = map(function(row){

    var prior = [.5,.5,.5]

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, prior,row.so)

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct_inf)

}, allData)

}

var steps = 5000
var samples = 1
var mll = AIS(model, {steps:steps, samples: samples})
mll
'
```

```{r}
# child_ex3_flat_prior_model_noise_likelihood<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.flat.prior.noise.marginal.likelihood , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   chains = 3,
#   cores = 3
# )%>%
#   saveRDS("../../stats/saves/child_ex3_flat_prior_model_noise_loglike.rds")

child_ex3_flat_prior_model_noise_likelihood <- readRDS("../../stats/saves/child_ex3_flat_prior_model_noise_loglike.rds")%>%
  as.tibble()%>%
  mutate(model = "flat_prior",
         parameter = "noise")%>%
  group_by(model,parameter)%>%
  summarise(logP = mean(V1))
```

##### Developmental noise parameter

```{r}
child.ex3.flat.prior.developmental.noise.marginal.likelihood <-'

var allData = dataFromR.allData

var model = function(){

  var noise_slope = uniformDrift({a: -5, b: 5, width: 0.2})
  var noise_int = uniformDrift({a: -5, b: 5, width: 0.1})

  var output = map(function(row){

    var noise = logistic(noise_int + noise_slope * (row.age_num-row.minage))

    var prior = [.5,.5,.5]

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, prior,row.so)

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct_inf)

}, allData)

}

var steps = 5000
var samples = 1
var mll = AIS(model, {steps:steps, samples: samples})
mll
'
```

```{r}
# child_ex3_flat_prior_model_developmental_noise_likelihood<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.flat.prior.developmental.noise.marginal.likelihood , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   chains = 3,
#   cores = 3
# )%>%
#   saveRDS("../../stats/saves/child_ex3_flat_prior_model_developmental_noise_loglike.rds")

child_ex3_flat_prior_model_developmental_noise_likelihood <- readRDS("../../stats/saves/child_ex3_flat_prior_model_developmental_noise_loglike.rds")%>%
  as.tibble()%>%
  mutate(model = "flat_prior",
         parameter = "developmental noise")%>%
  group_by(model,parameter)%>%
  summarise(logP = mean(V1))

```

#### Prior only model

##### Parameter free

```{r}
child.ex3.prior.only.marginal.likelihood <-'

var allData = dataFromR.allData

var model = function(){

  var noise = uniform({a: 0, b: 1})

  var output = map(function(row){

    var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: row.priorProbs}));
      return obj.shape == "circle" ? 1 : 0
       }})

    observe(modelPredictions, row.correct_inf)

}, allData)

}

var steps = 5000
var samples = 1
var mll = AIS(model, {steps:steps, samples: samples})
mll
'
```

```{r}
# child_ex3_prior_only_model_likelihood<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.prior.only.noise.marginal.likelihood , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   chains = 3,
#   cores = 3
# )%>%
#   saveRDS("../../stats/saves/child_ex3_prior_only_model_loglike.rds")

child_ex3_prior_only_model_likelihood <- readRDS("../../stats/saves/child_ex3_prior_only_model_loglike.rds")%>%
  as.tibble()%>%
  mutate(model = "prior_only",
         parameter = "parameter free")%>%
  group_by(model,parameter)%>%
  summarise(logP = mean(V1))
```

##### Noise parameter

```{r}
child.ex3.prior.only.noise.marginal.likelihood <-'

var allData = dataFromR.allData

var model = function(){

  var noise = uniform({a: 0, b: 1})

  var output = map(function(row){

    var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: row.priorProbs}));
      return obj.shape == "circle" ? 1 : 0
       }})

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct_inf)

}, allData)

}

var steps = 5000
var samples = 1
var mll = AIS(model, {steps:steps, samples: samples})
mll
'
```

```{r}
# child_ex3_prior_only_model_noise_likelihood<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.prior.only.noise.marginal.likelihood , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   chains = 3,
#   cores = 3
# )%>%
#   saveRDS("../../stats/saves/child_ex3_prior_only_model_noise_loglike.rds")

child_ex3_prior_only_model_noise_likelihood <- readRDS("../../stats/saves/child_ex3_prior_only_model_noise_loglike.rds")%>%
  as.tibble()%>%
  mutate(model = "prior_only",
         parameter = "noise")%>%
  group_by(model,parameter)%>%
  summarise(logP = mean(V1))
```

##### Developmental noise parameter

```{r}
child.ex3.prior.only.developmental.noise.marginal.likelihood <-'

var allData = dataFromR.allData

var model = function(){

  var noise_slope = uniformDrift({a: -5, b: 5, width: 0.2})
  var noise_int = uniformDrift({a: -5, b: 5, width: 0.1})

  var output = map(function(row){

    var noise = logistic(noise_int + noise_slope * (row.age_num-row.minage))

    var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: row.priorProbs}));
      return obj.shape == "circle" ? 1 : 0
       }})

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct_inf)

}, allData)

}

var steps = 5000
var samples = 1
var mll = AIS(model, {steps:steps, samples: samples})
mll
'
```

```{r}
# child_ex3_prior_only_model_developmental_noise_likelihood<- webppl(
#   program_code = paste(rsaUtils, rsaModel, child.ex3.prior.only.developmental.noise.marginal.likelihood , sep='\n'),
#   data =list(allData = ex3_model_data),
#   data_var = "dataFromR",
#   chains = 3,
#   cores = 3
# )%>%
#   saveRDS("../../stats/saves/child_ex3_prior_only_model_developmental_noise_loglike.rds")

child_ex3_prior_only_model_developmental_noise_likelihood <- readRDS("../../stats/saves/child_ex3_prior_only_model_developmental_noise_loglike.rds")%>%
  as.tibble()%>%
  mutate(model = "prior_only",
         parameter = "developmental noise")%>%
  group_by(model,parameter)%>%
  summarise(logP = mean(V1))
```

```{r model loglike ex3 children}
# calculating model log-likelihoods

# child_ex3_model_comparison <- bind_rows(
#   child_ex3_pragm_model_likelihood,
#   child_ex3_pragm_model_noise_likelihood,
#   child_ex3_pragm_model_developmental_noise_likelihood,
#   child_ex3_flat_prior_model_noise_likelihood,
#   child_ex3_flat_prior_model_likelihood,
#   child_ex3_flat_prior_model_developmental_noise_likelihood,
#   child_ex3_prior_only_model_noise_likelihood,
#   child_ex3_prior_only_model_likelihood,
#   child_ex3_prior_only_model_developmental_noise_likelihood
# )%>%
#   ungroup()%>%
#   mutate(model = reorder(model,logP))%>%
#   saveRDS("../../stats/saves/child_ex3_model_comparison.rds")

child_ex3_model_comparison <- readRDS("../../stats/saves/child_ex3_model_comparison.rds")

```

### Plot for likelihoods

```{r plot model loglikes ex3 children, include = T}
ggplot(child_ex3_model_comparison, aes(x = model, y = logP, col = model))+
  geom_segment( aes(x=model, xend=model, y=-400, yend=logP), col = "black", size = 1)+
  geom_point(size = 7)+
  theme_few()+ 
  xlab("Model")+
  ylab("Log-likelihood")+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  guides(col = F)+
  facet_grid(~parameter)+
  ylim(-475,-400)+
  scale_color_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#859900","#b58900","#6c71c4"))

```

### Bayes factors

```{r, include = T}
child_ex3_pragmatic_models_bf <- child_ex3_model_comparison%>%
  filter(model == "pragmatic")%>%
  spread(parameter, logP) %>%
  mutate("dev. noise > noise" = exp(`developmental noise` - noise),
         "noise > parameter free" = exp(noise - `parameter free`),
         "dev. noise > parameter free" = exp(`developmental noise` - `parameter free`)) %>%
  select(-`parameter free`,-`developmental noise`, -noise)

child_ex3_pragmatic_models_bf%>%
  kable(digits = 2)
```

```{r bayes factors ex3 children, include = T}
child_ex3_bf <- child_ex3_model_comparison%>%
  group_by(parameter)%>%
  spread(model, logP) %>%
  mutate("pragmatic > flat_prior" = pragmatic - flat_prior,
         "pragmatic > prior_only" = pragmatic - prior_only,
         "flat_prior > prior_only" = flat_prior - prior_only) %>%
  select(-pragmatic,-flat_prior, -prior_only)

child_ex3_bf%>%
  kable(digits = 2)
```